---
title: "Total_soil_c_comparisons"
author: "Becca"
date: "5/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Comparing bulk organic C values (tons/hectare) for soil depths 1 and 2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in data and libraries
```{r }
#install.packages("mice")
library(mice)
library(plyr)
library(dplyr)

hhout <- read.csv("hhout_master_01May2020.csv", header=TRUE)
summary(hhout)
hhout$depth <-as.factor(hhout$depth)
summary(hhout$bulk.dens)
#there are missing bulk density values but they represent a small number of sites and are interspersed among species, so we'll try to do an analysis without the missing data.
```


Calculate bulk soil organic C tons per hectare 
```{r }
#first make a cm depth variable
hhout$depth.cm <- revalue(hhout$depth, c("1"="5", "2"="10", "3"="10", "4"="10"))
#summary(hhout$depth)
#summary(hhout$depth.cm)
hhout$depth.cm <- as.numeric(as.character(hhout$depth.cm))
summary(hhout$depth.cm)

#calculate the tons per hectare
hhout$tons.c.hectare <- (hhout$org.c.in.total.bulk * hhout$bulk.dens * hhout$depth.cm)
summary(hhout$tons.c.hectare)
```


See whether species correlates with tons c per hectare
```{r }
totalc.lm <- aov(tons.c.hectare ~ species*depth, data = hhout)
summary(totalc.lm)
TukeyHSD(totalc.lm)
```

Fagus is significantly higher than all of the other species in depth 1 but there is no difference among species at depth 2. 

See whether litter chemistry is associated with total soil C at depth 1
```{r }
hhout.depth1 <- subset(hhout, depth==1)

totalc.litter.lm.1 <- lm(tons.c.hectare ~ litter.pct.c + litter.pct.n + litter.lignin.mg.g, data = hhout.depth1)
summary(totalc.litter.lm.1)

totalc.litter.interaction.lm.1 <- lm(tons.c.hectare ~ litter.pct.c + litter.pct.n*litter.lignin.mg.g, data = hhout.depth1)
summary(totalc.litter.interaction.lm.1)
```


See whether litter chemistry is associated with total soil C at depth 2
```{r }
hhout.depth2 <- subset(hhout, depth==2)

totalc.litter.lm.2 <- lm(tons.c.hectare ~ litter.pct.c + litter.pct.n + litter.lignin.mg.g, data = hhout.depth2)
summary(totalc.litter.lm.2)

totalc.litter.interaction.lm.2 <- lm(tons.c.hectare ~ litter.pct.c + litter.pct.n*litter.lignin.mg.g, data = hhout.depth2)
summary(totalc.litter.interaction.lm.2)

AIC(totalc.litter.lm.2, totalc.litter.interaction.lm.2)
```
The model without the interaction is better than the one with the interaction. 


Is soil pH or moisture associated with total C at depth 1?
```{r }
totalc.soil.lm.1 <- lm(tons.c.hectare ~ soil.ph + pct.m, data = hhout.depth1)
summary(totalc.soil.lm.1)
```` 

Soil pH has a negative relationship (higher pH = less C) at depth 1

Is soil pH or moisture associated with total C at depth 2?
```{r }
totalc.soil.lm.2 <- lm(tons.c.hectare ~ soil.ph + pct.m, data = hhout.depth2)
summary(totalc.soil.lm.2)
```` 

pH still has a negative correlation but it is not significant. 