---
title: "Total_soil_c_comparisons"
author: "Becca"
date: "5/1/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Comparing bulk organic C values (tons/hectare) for soil depths 1 and 2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in data and libraries
```{r }
#install.packages("mice")
library(tidyverse)
library(stats)
library(dplyr)

hhout <- read.csv("hhout_master_01May2020.csv", header=TRUE)
summary(hhout)
hhout$depth <-as.factor(hhout$depth)
summary(hhout$bulk.dens)
#there are missing bulk density values but they represent a small number of sites and are interspersed among species, so we'll try to do an analysis without the missing data.
```


Calculate bulk soil organic C tons per hectare 
```{r }
#first make a cm depth variable
hhout$depth.cm <- revalue(hhout$depth, c("1"="5", "2"="10", "3"="10", "4"="10"))
#summary(hhout$depth)
#summary(hhout$depth.cm)
hhout$depth.cm <- as.numeric(as.character(hhout$depth.cm))
summary(hhout$depth.cm)

#there is a mistake in the bulk density
hhout[hhout$bulk.dens>2, ]
#row 22, bulk density has been given same value as pH. Remove this value of bulk density
hhout$bulk.dens[c(64:66)] <- NA

#calculate the tons organic C per hectare
hhout$tons.c.hectare <- hhout$org.c.in.total.bulk * hhout$bulk.dens * hhout$depth.cm
summary(hhout$tons.c.hectare)
```


Summarize bulk soil data for paper
```{r}
#first remove the fraction codes and select the unique values of bulk soil
bulk_soil <- hhout[ , c("site", "species", "depth", "tons.c.hectare", "pct.n.pre.ashing.bulk", "pct.c.pre.ashing.bulk", "pct.n.in.org.fraction.bulk", "pct.c.in.org.fraction.bulk", "org.c.start.mass.bulk", "soil.ph", "bulk.dens", "pct.m", "soil.doc.mass.g")]

#bulk_soil <- na.omit(bulk_soil)
bulk_soil <- distinct(bulk_soil)
summary(bulk_soil)


#summarize the data frame
bulk_summary <- bulk_soil%>% 
  dplyr::group_by(species, depth)%>% 
  dplyr::summarize(
    tons.c = mean(tons.c.hectare, na.rm = TRUE),
    sd_tons.c = sd(tons.c.hectare, na.rm = TRUE),
    pct.n.pre = mean(pct.n.pre.ashing.bulk, na.rm=TRUE),
    pct.c.pre = mean(pct.c.pre.ashing.bulk, na.rm=TRUE),
    pct.n.org = mean(pct.n.in.org.fraction.bulk, na.rm=TRUE), 
    pct.c.org = mean(pct.c.in.org.fraction.bulk, na.rm=TRUE),
    ph = mean(soil.ph, na.rm=TRUE), 
    sd_ph = sd(soil.ph, na.rm=TRUE),
    bulkdens = mean(bulk.dens, na.rm=TRUE),
    sd_bd = sd(bulk.dens, na.rm=TRUE),
    moist = mean(pct.m, na.rm=TRUE), 
    sd_moist = sd(pct.m, na.rm=TRUE),
    sd.n.pre = sd(pct.n.pre.ashing.bulk, na.rm=TRUE),
    sd.org.n = sd(pct.n.in.org.fraction.bulk, na.rm=TRUE), 
    sd.c.pre = sd(pct.c.pre.ashing.bulk, na.rm=TRUE), 
    sd.org.c = sd(pct.c.in.org.fraction.bulk, na.rm=TRUE))
print(bulk_summary, row.names = FALSE)
write.csv(bulk_summary, file = "bulk.summary.csv", row.names = FALSE)
```


See whether species correlates with tons c per hectare
```{r }
totalc.lm <- aov(tons.c.hectare ~ species*depth, data = hhout)
summary(totalc.lm)
TukeyHSD(totalc.lm)
```

Fagus is significantly higher than all of the other species in depth 1 but there is no difference among species at depth 2. 

See whether litter chemistry is associated with total soil C at depth 1
```{r }
hhout.depth1 <- subset(hhout, depth==1)

totalc.litter.lm.1 <- lm(tons.c.hectare ~ litter.pct.c + litter.pct.n + litter.lignin.mg.g, data = hhout.depth1)
summary(totalc.litter.lm.1)

totalc.litter.interaction.lm.1 <- lm(tons.c.hectare ~ litter.pct.c + litter.pct.n*litter.lignin.mg.g, data = hhout.depth1)
summary(totalc.litter.interaction.lm.1)

AIC(totalc.litter.lm.1, totalc.litter.interaction.lm.1)
```
Almost identical in explanatory value. 

See whether litter chemistry is associated with total soil C at depth 2
```{r }
hhout.depth2 <- subset(hhout, depth==2)

totalc.litter.lm.2 <- lm(tons.c.hectare ~ litter.pct.c + litter.pct.n + litter.lignin.mg.g, data = hhout.depth2)
summary(totalc.litter.lm.2)

totalc.litter.interaction.lm.2 <- lm(tons.c.hectare ~ litter.pct.c + litter.pct.n*litter.lignin.mg.g, data = hhout.depth2)
summary(totalc.litter.interaction.lm.2)

AIC(totalc.litter.lm.2, totalc.litter.interaction.lm.2)
```
The model without the interaction is better than the one with the interaction. 


Is soil pH or moisture associated with total C at depth 1?
```{r }
totalc.soil.lm.1 <- lm(tons.c.hectare ~ soil.ph + pct.m, data = hhout.depth1)
summary(totalc.soil.lm.1)
```` 

Soil pH has a negative relationship (higher pH = less C) at depth 1

Is soil pH or moisture associated with total C at depth 2?
```{r }
totalc.soil.lm.2 <- lm(tons.c.hectare ~ soil.ph + pct.m, data = hhout.depth2)
summary(totalc.soil.lm.2)
```` 

pH still has a negative correlation but it is not significant. 