---
title: "CN_Ratios"
author: "Becca"
date: "11-7-2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Script to reproduce CN ratios of soil samples (fractions for depths 1 and 2, bulk for depths 3-4)

###Data and libraries
```{r }
library(tidyverse)
library(dplyr)
library(ggplot2)
cn.wts.sums <- read.csv("cn.wts.sums.csv", header=TRUE)
#correct read-in errors
summary(cn.wts.sums)
cn.wts.sums <- cn.wts.sums[,-c(1,2)]
cn.wts.sums$depth = as.factor(cn.wts.sums$depth)
#check dataframe again
summary(cn.wts.sums)
```


###CN ratios

Check the CN ratios for pre and post-ashing bulk samples. 
```{r }
cn.ratios.bulk <- cn.wts.sums

cn.ratios.bulk$cn.pre.bulk <- cn.ratios.bulk$mean.c.pre/cn.ratios.bulk$mean.n.pre  
cn.ratios.bulk$cn.post.bulk <- cn.ratios.bulk$mean.c.post/cn.ratios.bulk$mean.n.post

(ratioplot.pre<-ggplot(cn.ratios.bulk, aes(x=species, y=cn.pre.bulk)) +
    geom_point() +
    facet_wrap(~depth))
ratioplot.pre

ratioplot.post<-ggplot(cn.ratios.bulk, aes(x=species, y=cn.post.bulk)) + 
    geom_point() +
    facet_wrap(~depth)
ratioplot.post
```

CN ratios pre-ashing bulk appear in the normal range for soil. Post-ashing, they are extremely high, likely because nearly all of the N was lost to volatilization during combustion. 

**Now check CN ratios for each fraction pre-and-post ashing**

```{r }
#make the pom dataframe
cn.ratios.pom <- subset(cn.wts.sums, fraction.code %in% 'pom')
    cn.ratios.pom$cn.pre = (cn.ratios.pom$pct.c.pre.ashing/cn.ratios.pom$pct.n.pre.ashing) 
    cn.ratios.pom$cn.post = (cn.ratios.pom$pct.c.post.ashing/cn.ratios.pom$pct.n.post.ashing) 
    cn.ratios.pom$cn.org = (cn.ratios.pom$org.c.in.t.frac/cn.ratios.pom$org.n.in.t.frac)
```


Plot the cn ratios for the pom pre, post, and organic
```{r }
(ratioplot.pre.pom<-ggplot(cn.ratios.pom, aes(x=species, y=cn.pre)) +
    geom_point() +
    facet_wrap(~depth))
ratioplot.pre.pom

ratioplot.post.pom<-ggplot(cn.ratios.pom, aes(x=species, y=cn.post)) + 
    geom_point() +
    facet_wrap(~depth)
ratioplot.post.pom

ratioplot.organic.pom <- ggplot(cn.ratios.pom, aes(x=species, y=cn.org)) +
    geom_point() +
    facet_wrap(~depth)
ratioplot.organic.pom
```


Make the sa dataframe
```{r results="hide"}
cn.ratios.sa <- subset(cn.wts.sums, fraction.code %in% 'sa')
    cn.ratios.sa$cn.pre = (cn.ratios.sa$pct.c.pre.ashing/cn.ratios.sa$pct.n.pre.ashing) 
    cn.ratios.sa$cn.post = (cn.ratios.sa$pct.c.post.ashing/cn.ratios.sa$pct.n.post.ashing) 
    cn.ratios.sa$cn.org = (cn.ratios.sa$org.c.in.t.frac/cn.ratios.sa$org.n.in.t.frac)
```

Plot the cn ratios for the sa pre, post, and organic
```{r }
ratioplot.pre.sa<-ggplot(cn.ratios.sa, aes(x=species, y=cn.pre)) +
    geom_point() +
    facet_wrap(~depth)
ratioplot.pre.sa

ratioplot.post.sa<-ggplot(cn.ratios.sa, aes(x=species, y=cn.post)) + 
    geom_point() +
    facet_wrap(~depth)
ratioplot.post.sa

ratioplot.organic.sa <- ggplot(cn.ratios.sa, aes(x=species, y=cn.org)) +
    geom_point() +
    facet_wrap(~depth)
ratioplot.organic.sa
```


Make the sc dataframe
```{r results="hide"}
cn.ratios.sc <- subset(cn.wts.sums, fraction.code %in% 'sc')
    cn.ratios.sc$cn.pre = (cn.ratios.sc$pct.c.pre.ashing/cn.ratios.sc$pct.n.pre.ashing) 
    cn.ratios.sc$cn.post = (cn.ratios.sc$pct.c.post.ashing/cn.ratios.sc$pct.n.post.ashing) 
    cn.ratios.sc$cn.org = (cn.ratios.sc$org.c.in.t.frac/cn.ratios.sc$org.n.in.t.frac)
```


Plot the cn ratios for the sc pre, post, and organic
```{r }
ratioplot.pre.sc<-ggplot(cn.ratios.sc, aes(x=species, y=cn.pre)) +
    geom_point() +
    facet_wrap(~depth)
ratioplot.pre.sc

ratioplot.post.sc<-ggplot(cn.ratios.sc, aes(x=species, y=cn.post)) + 
    geom_point() +
    facet_wrap(~depth)
ratioplot.post.sc

ratioplot.organic.sc <- ggplot(cn.ratios.sc, aes(x=species, y=cn.org)) +
    geom_point() +
    facet_wrap(~depth)
ratioplot.organic.sc
```

**The CN ratios are quite consistend despite large variation in actual values. Maybe this means that the problem isn't in the measurements but somewhere else??**

