---
title: "CUE_for_paper"
author: "Becca"
date: "3/18/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Doing the microbial growth rate and CUE analysis with the master data frame and the cue data frame. 

Load libraries
```{r }
library(ggplot2)
library(plyr)
library(tidyverse)
#library(plyr)
#library(dplyr)
library(AICcmodavg)
library(FlexDir)
library(DirichletReg)
```


Load CUE data and make the species levels match with the master dataframe
```{ r }
cue <- read.csv("field_CUE_17.csv", header=TRUE)
summary(cue)

#outlier in row 33
cue_corrected<-cue[-33, ]
summary(cue_corrected)

#combine picea sitchensis and picea abies into one group
cue_corrected$species<-mapvalues(cue_corrected$species, from=c("Acer", "Fagus", "Fraxinus", "Picea abies", "Picea sitchensis", "Populus", "Quercus"), to=c("acer", "fagus", "fraxinus", "picea", "picea", "populus", "quercus"))
#check that it worked
levels(cue_corrected$species)
#looks good!
```


Read in the master dataframe
```{ r }
hhout <- read.csv("hhout_master", header=TRUE)
summary(hhout)

#make depth a factor
hhout$depth <- as.factor(hhout$depth)

#subset out just depth 1 (to match with CUE data)
hhout_sub <- hhout[hhout$depth=="1", ]
summary(hhout_sub)  
```


Add CUE to the hhout_sub data
```{ r }
hhout_sub_cue <- join(hhout_sub, cue_corrected, type="left")
summary(hhout_sub_cue)

#rename for analysis convenience
hh_cue <- hhout_sub_cue

#write the data frame for convenience
write.csv(hh_cue, file = "hh_cue.csv", row.names = FALSE)
```


Plot CUE as function of different soil and litter parameters
```{r }
#read in the combined dataframe
# hh_cue <- read.csv("hh_cue.csv", header=TRUE)
# summary(hh_cue)
# hh_cue$depth = as.factor(hh_cue$depth)

#calculate HWE-C as mg C per g dry weight litter
# hhout$npoc.mg.g <- (hhout$soil.npoc.mg.L * 0.04) / 5
# summary(hhout$npoc.mg.g)
# hh_cue$litter.hwec.mg.g <- (hh_cue$litter.hwec.ppm * 0.2 / 2)
# summary(hh_cue$litter.hwec.mg.g)
# summary(hh_cue$litter.cell.mg.g)

#write.csv(hh_cue, file = "hh_cue.csv", row.names = FALSE)

hh_cue <- read.csv("hh_cue.csv", header=TRUE)
hh_cue %>% distinct() -> hh_cue


#HWE-c
(cueplot1 <- ggplot(hh_cue, aes(x = litter.hwec.mg.g, y = CUE)) +
              geom_point()+
              geom_smooth())

#soil pH
(cueplot2 <- ggplot(hh_cue, aes(x = soil.ph, y = CUE)) + 
              geom_point())

#litter lignin.n.ratio
#first make the ln ratio variable
hh_cue$lnratio <- (hh_cue$litter.lignin.mg.g/10)/hh_cue$litter.pct.n
#now make the plot
(cueplot3 <- ggplot(hh_cue, aes(x = lnratio, y=CUE)) + 
              geom_point() + geom_jitter() + geom_smooth(method="lm")) 


#try as function of latitude
(cueplot4 <- ggplot(hh_cue, aes(x = lat, y=CUE)) +
              geom_point())
```

There's just a huge range in CUE, doesn't seem to have a pattern. 

Make a model set based on potentially-important parameters.

**Basic parameters**
* species 

**Soil organic and inorganic carbon and nitrogen**
* pct.org.c.in.whole.bulk
* pct.org.n.in.whole.bulk
* pct.c.post.ashing.bulk
* pct.n.post.ashing.bulk
* soil.npoc.mg.L

**Soil moisture and ph**
* soil.ph
* pct.m

**litter chemistry**
* litter.pct.n
* litter.pct.c
* litter.lignin.mg.g
* litter.cell.mg.g
* litter.hwec.mg.g
* lnratio

First check which parameters are closely corerlated
```{ r}
library(GGally)

#subset the relevant parameters
cue_testing <- subset(hh_cue, select = c(species, pct.c.in.org.fraction.bulk, pct.n.in.org.fraction.bulk, pct.c.post.ashing.bulk, pct.n.post.ashing.bulk, soil.npoc.mg.L, soil.ph, pct.m, litter.pct.c, litter.pct.n, litter.lignin.mg.g, litter.cell.mg.g, litter.hwec.ppm, lnratio))
  
ggpairs(cue_testing)
```

**Pairs that are strongly related*:**
* +0.843: pct.c.in.org.fraction.bulk + pct.n.in.org.fraction.bulk

* +0.877: litter.lignin.mg.g + litter.pct.c
* -0.912: litter.lignin.mg.g + litter.pct.n

* +0.816: lnratio + litter.pct.c
* -0.957: lnratio + litter.pct.n
* +0.943: lnratio + litter.lignin.mg.g
* +0.804: lnratio + litter.cell.mg.g
* +0.804: lnratio + litter.cell.mg.g
* -0.933: lnratio + litter.hwec.ppm   

* -0.968: litter.hwec.ppm + litter.lignin.mg.g
* -0.692: litter.hwec.ppm + litter.cell.mg.g

* -0.773: litter.pct.n + litter.pct.c

* -0.938: litter.pct.c + litter.hwec.ppm
* -0.884: litter.hwec.ppm + litter.pct.c

* -0.840: litter.cell.mg.g + litter.pct.n
* +0.826: litter.cell.mg.g + litter.lignin.mg.g


**Variables to use out of the strongly-related variables**
* lnratio <- use in place of litter.hwec.mg.g (or if you really want use hwec.ppm and leave out lnratio)
* lnratio <- use in place of litter.pct.c, litter.cell.mg.g, litter.pct.n, litter.lignin.mg.g
* litter.pct.c <- use in place of litter.hwec.ppm (or make a pct.c:hwec ratio)
* pct.org.c.in.whole.bulk <- use in place of pct.org.n.in.whole.bulk

**Variables to use from the non-strongly-related variables**
* soil ph
* soil.pct.m - don't use this one, use the pct.m from the CUE calculations
* species (maybe...but probably not)
* soil.doc.mass.g
* inorganic c or n - no reason to think this is important - maybe try %n post ashing

Make a litter.ccratio variable for ratio between litter c and litter hwe-c
```{r }
#first 
hh_cue$litter.ccratio <- hh_cue$litter.pct.c/(hh_cue$litter.hwec.mg.g/10)
summary(hh_cue$litter.ccratio)
``` 


Make the model set
```{r }
#single models
cue.null <- lm(CUE ~ 1, data = hh_cue)
cue.lnratio <- lm(CUE ~ lnratio, data = hh_cue)
cue.litter.c <- lm(CUE ~ litter.pct.c, data = hh_cue)
cue.litter.cc <- lm(CUE ~ litter.ccratio, data = hh_cue)
cue.soil.ph <- lm(CUE ~ soil.ph, data=hh_cue)
cue.npoc <- lm(CUE ~ soil.doc.mass.g, data = hh_cue)
cue.inorg.n <- lm(CUE ~ pct.n.post.ashing.bulk, data = hh_cue)
cue.org.c <- lm(CUE ~ org.c.in.total.bulk, data = hh_cue)
cue.

#2-level models 
# cue.ln.ph <- lm(CUE ~ lnratio + soil.ph, data = hh_cue)
# cue.ln.npoc <- lm(CUE ~ lnratio + soil.doc.mass.g, data = hh_cue)
# cue.ln.inorg.n <- lm(CUE ~ lnratio + pct.n.post.ashing.bulk, data=hh_cue)
# cue.ln.orgc <- lm(CUE ~ lnratio + org.c.in.total.bulk, data = hh_cue)
# cue.cc.ph <- lm(CUE ~ litter.ccratio + soil.ph, data = hh_cue)
# cue.cc.npoc <- lm(CUE ~ litter.ccratio + soil.doc.mass.g, data = hh_cue)
# cue.cc.inorg.n <- lm(CUE ~ litter.ccratio + pct.n.post.ashing.bulk, data = hh_cue)
# cue.cc.orgc <- lm(CUE ~ litter.ccratio + org.c.in.total.bulk, data = hh_cue)

#this model set is redundant, try combining some


#interaction models
``` 


Compare models
*turned off this chunk because these models were mostly redundant
```{ r}
model.set <- list(cue.null, cue.lnratio, cue.litter.c, cue.litter.cc, cue.soil.ph, cue.npoc, cue.inorg.n, cue.org.c, cue.ln.ph, cue.ln.npoc, cue.ln.inorg.n, cue.ln.orgc, cue.cc.ph, cue.cc.npoc, cue.cc.inorg.n, cue.inorg.n, cue.cc.orgc)

modnames <- c("cue.null", "cue.lnratio", "cue.litter.c", "cue.litter.cc", "cue.soil.ph", "cue.npoc", "cue.inorg.n", "cue.org.c", "cue.ln.ph", "cue.ln.npoc", "cue.ln.inorg.n", "cue.ln.orgc", "cue.cc.ph", "cue.cc.npoc", "cue.cc.inorg.n", "cue.inorg.n", "cue.cc.orgc")

(cue.compare <- aictab(cand.set = model.set, modnames = modnames))
``` 


compare just the single models
```{r}
single.model.set <- list(cue.null, cue.lnratio, cue.litter.c, cue.litter.cc, cue.soil.ph, cue.npoc, cue.inorg.n, cue.org.c)

single.modnames <- c("cue.null", "cue.lnratio", "cue.litter.c", "cue.litter.cc", "cue.soil.ph", "cue.npoc", "cue.inorg.n", "cue.org.c")

(cue.compare <- aictab(cand.set = single.model.set, modnames = single.modnames))

```

Litter carbon carries 91% of the model weight, but the next-best models are still within delta-2 AIC: 

Model selection based on AICc:

              K    AICc Delta_AICc AICcWt Cum.Wt    LL
cue.litter.c  3 -102.16       0.00   0.43   0.43 54.24
cue.npoc      3 -101.04       1.11   0.25   0.68 53.68
cue.inorg.n   3 -100.99       1.16   0.24   0.92 53.66
cue.litter.cc 3  -98.06       4.10   0.06   0.97 52.19
cue.null      2  -94.89       7.26   0.01   0.98 49.53
cue.lnratio   3  -94.82       7.34   0.01   0.99 50.57
cue.soil.ph   3  -92.73       9.43   0.00   1.00 49.54
cue.org.c     3  -90.55      11.61   0.00   1.00 48.44

``` {r}
summary(cue.litter.c)
#A small but significant increase in CUE (with increasing litter C (r2 = 0.05, p = 0.002).

summary(cue.npoc)
#small but significant decline in CUE with increasing DOC in the topsoil (r2 = -143.76, p = 0.001), but this relationship seems to be driven by a few outliers.

summary(cue.inorg.n)
#small but significant decrease in CUE with increasing inorganic N ()
``` 


Plot the data on a graph:
```{r }
(cue.litterc.plot <- ggplot(hh_cue, aes(x = litter.pct.c, y = CUE)) + 
                              geom_point() +
                              geom_smooth(method = "lm", color = " black"))

(doc.cue.plot <- ggplot(hh_cue, aes(x = soil.doc.mass.g, y = CUE)) + 
                              geom_point() +
                              geom_smooth(method = "lm", color = " black"))

(inorg.n.cue.plot <- ggplot(hh_cue, aes(x = pct.n.post.ashing.bulk, y = CUE)) + 
                              geom_point() +
                              geom_smooth(method = "lm", color = " black"))
``` 


Significant variation by litter cc ratio, but a way worse model than model 1 - only explains about half as much variation. Regardless, both have fairly low predictive power. 
```{r }
(cue.litter.cc.plot <- ggplot(hh_cue, aes(x = litter.ccratio, y = CUE)) + 
                              geom_point() +
                              geom_smooth(method = "lm", color = " black"))
``` 


Test the relationship between microbial cue and soil C distribution

First get the normalized diri distribution
```{r }
names(hhout)


levels(hhout$fractionated)
#select out just the data with fraction info
fraction_data <- hhout[hhout$fractionated=="y",]
summary(fraction_data)

#subset out just relevant variables
fraction_data <- select(hhout, c(site, species, depth, fraction.code, org.c.mass.fraction))
fraction_data <- fraction_data[complete.cases(fraction_data),]
summary(fraction_data)

#spread the values of fraction mass over fraction code
fraction_wide <- spread(fraction_data, fraction.code, org.c.mass.fraction)
fraction_wide <- fraction_wide[complete.cases(fraction_wide),]

names(fraction_wide)
summary(fraction_wide)
#Make the FD normalized data


fraction.diri <- FD.normalization(fraction_wide[,(4:6)])
fraction.diri


#bind the fraction data to the fraction diri
fraction_wide <- cbind(fraction_wide, fraction.diri)
summary(fraction_wide)

fraction_wide <- fraction_wide[, -c(4:6)]
head(fraction_wide)
```


Join the cue to the diri df
```{r}
fraction_wide_unique <- fraction_wide[unique(fraction_wide$site), ]
summary(fraction_wide_unique)

summary(cue_corrected)
cue_corrected_drop <- cue_corrected[,-1]

#join the cue data to the fraction wide data
fraction_wide_cue <- join(fraction_wide_unique, cue_corrected_drop, type = "left")
#check the join
summary(fraction_wide_cue)
length(fraction_wide_cue$site)
fraction_wide_cue
``` 


Now do the correlation test
```{r }
orgs <- DR_data(fraction_wide_cue[,4:6])

cue.mod <- DirichReg(orgs ~ CUE + depth, data = fraction_wide_cue)
summary(cue.mod)
``` 

Microbial CUE is not related to soil organic carbon distribution among fractions. 

This suggests that both microbial CUE and soc distribution are responses to litter chemistry.  







