---
title: "CUE_for_paper"
author: "Becca"
date: "3/18/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Doing the microbial growth rate and CUE analysis with the master data frame and the cue data frame. 

Load libraries
```{r }
library(ggplot2)
library(plyr)
library(tidyverse)
#library(plyr)
#library(dplyr)
library(AICcmodavg)
library(FlexDir)
library(DirichletReg)
```


Load CUE data and make the species levels match with the master dataframe
```{ r }
cue <- read.csv("field_CUE_17.csv", header=TRUE)
summary(cue)

#outlier in row 33
cue_corrected<-cue[-33, ]
summary(cue_corrected)

#combine picea sitchensis and picea abies into one group
cue_corrected$species<-mapvalues(cue_corrected$species, from=c("Acer", "Fagus", "Fraxinus", "Picea abies", "Picea sitchensis", "Populus", "Quercus"), to=c("acer", "fagus", "fraxinus", "picea", "picea", "populus", "quercus"))
#check that it worked
levels(cue_corrected$species)
#looks good!
```


Read in the master dataframe
```{ r }
hhout <- read.csv("hhout_master_01May2020.csv", header=TRUE)
summary(hhout)

#make depth a factor
hhout$depth <- as.factor(hhout$depth)

#subset out just depth 1 (to match with CUE data)
hhout_sub <- hhout[hhout$depth=="1", ]
summary(hhout_sub)  
```


Add CUE to the hhout_sub data
```{ r }
hhout_sub_cue <- join(hhout_sub, cue_corrected, type="left")
summary(hhout_sub_cue)

#rename for analysis convenience
hh_cue <- hhout_sub_cue

#write the data frame for convenience
write.csv(hh_cue, file = "hh_cue.csv", row.names = FALSE)
```


Plot CUE as function of different soil and litter parameters
```{ r }
#read in the combined dataframe
hh_cue <- read.csv("hh_cue.csv", header=TRUE)
summary(hh_cue)
hh_cue$depth = as.factor(hh_cue$depth)

#calculate HWE-C as mg C per g dry weight litter
hhout$npoc.mg.g <- (hhout$soil.npoc.mg.L * 0.04) / 5
summary(hhout$npoc.mg.g)
hh_cue$litter.hwec.mg.g <- (hh_cue$litter.hwec.ppm * 0.2 / 2)
summary(hh_cue$litter.hwec.mg.g)
summary(hh_cue$litter.cell.mg.g)

write.csv(hh_cue, file = "hh_cue.csv", row.names = FALSE)

hh_cue <- read.csv("hh_cue.csv", header=TRUE)
hh_cue %>% distinct() -> hh_cue


#HWE-c
(cueplot1 <- ggplot(hh_cue, aes(x = litter.hwec.mg.g, y = CUE)) +
              geom_point()+
              geom_smooth(method="lm"))

#soil pH
(cueplot2 <- ggplot(hh_cue, aes(x = soil.ph, y = CUE)) + 
              geom_point())

#litter lignin.n.ratio
#first make the ln ratio variable
hh_cue$lnratio <- (hh_cue$litter.lignin.mg.g/10)/hh_cue$litter.pct.n
#now make the plot
(cueplot3 <- ggplot(hh_cue, aes(x = lnratio, y=CUE)) + 
              geom_point() + geom_jitter() + geom_smooth(method="lm")) 


#try as function of latitude
(cueplot4 <- ggplot(hh_cue, aes(x = lat, y=CUE)) +
              geom_point())
```

There's just a huge range in CUE, doesn't seem to have a pattern. 

Make a model set based on potentially-important parameters.

**Basic parameters**
* species 

**Soil organic and inorganic carbon and nitrogen**
* pct.org.c.in.whole.bulk
* pct.org.n.in.whole.bulk
* pct.c.post.ashing.bulk
* pct.n.post.ashing.bulk
* soil.npoc.mg.L

**Soil moisture and ph**
* soil.ph
* pct.m

**litter chemistry**
* litter.pct.n
* litter.pct.c
* litter.lignin.mg.g
* litter.cell.mg.g
* litter.hwec.mg.g
* lnratio

First check which parameters are closely corerlated
```{ r}
library(GGally)

#subset the relevant parameters
cue_testing <- subset(hh_cue, select = c(species, pct.c.in.org.fraction.bulk, pct.n.in.org.fraction.bulk, pct.c.post.ashing.bulk, pct.n.post.ashing.bulk, soil.npoc.mg.L, soil.ph, pct.m, litter.pct.c, litter.pct.n, litter.lignin.mg.g, litter.cell.mg.g, litter.hwec.ppm, lnratio))
  
ggpairs(cue_testing)
```

**Pairs that are strongly related*:**
* +0.843: pct.c.in.org.fraction.bulk + pct.n.in.org.fraction.bulk

* +0.877: litter.lignin.mg.g + litter.pct.c
* -0.912: litter.lignin.mg.g + litter.pct.n

* +0.816: lnratio + litter.pct.c
* -0.957: lnratio + litter.pct.n
* +0.943: lnratio + litter.lignin.mg.g
* +0.804: lnratio + litter.cell.mg.g
* +0.804: lnratio + litter.cell.mg.g
* -0.933: lnratio + litter.hwec.ppm   

* -0.968: litter.hwec.ppm + litter.lignin.mg.g
* -0.692: litter.hwec.ppm + litter.cell.mg.g

* -0.773: litter.pct.n + litter.pct.c

* -0.938: litter.pct.c + litter.hwec.ppm
* -0.884: litter.hwec.ppm + litter.pct.c

* -0.840: litter.cell.mg.g + litter.pct.n
* +0.826: litter.cell.mg.g + litter.lignin.mg.g


**Variables to use out of the strongly-related variables**
* lnratio <- use in place of litter.hwec.mg.g (or if you really want use hwec.ppm and leave out lnratio)
* lnratio <- use in place of litter.pct.c, litter.cell.mg.g, litter.pct.n, litter.lignin.mg.g
* litter.pct.c <- use in place of litter.hwec.ppm (or make a pct.c:hwec ratio)
* pct.org.c.in.whole.bulk <- use in place of pct.org.n.in.whole.bulk

**Variables to use from the non-strongly-related variables**
* soil ph
* soil.pct.m - don't use this one, use the pct.m from the CUE calculations
* species (maybe...but probably not)
* soil.doc.mass.g
* inorganic c or n - no reason to think this is important - maybe try %n post ashing

Make a litter.ccratio variable for ratio between litter c and litter hwe-c
```{r }
#first 
hh_cue$litter.ccratio <- hh_cue$litter.pct.c/(hh_cue$litter.hwec.mg.g/10)
summary(hh_cue$litter.ccratio)
``` 


Make the model set
```{r }
#single models
cue.null <- lm(CUE ~ 1, data = hh_cue)
cue.lnratio <- lm(CUE ~ lnratio, data = hh_cue)
cue.litter.c <- lm(CUE ~ litter.pct.c, data = hh_cue)
cue.litter.cc <- lm(CUE ~ litter.ccratio, data = hh_cue)
cue.soil.ph <- lm(CUE ~ soil.ph, data=hh_cue)
cue.doc <- lm(CUE ~ soil.doc.mass.g, data = hh_cue)
cue.inorg.n <- lm(CUE ~ pct.n.post.ashing.bulk, data = hh_cue)
cue.org.c <- lm(CUE ~ org.c.in.total.bulk, data = hh_cue)
cue.litter.lignin <- lm(CUE ~ litter.lignin.mg.g, data = hh_cue)
cue.org.n <- lm(CUE ~ pct.n.in.total.bulk, data = hh_cue)

#2-level models 
# cue.ln.ph <- lm(CUE ~ lnratio + soil.ph, data = hh_cue)
# cue.ln.npoc <- lm(CUE ~ lnratio + soil.doc.mass.g, data = hh_cue)
# cue.ln.inorg.n <- lm(CUE ~ lnratio + pct.n.post.ashing.bulk, data=hh_cue)
# cue.ln.orgc <- lm(CUE ~ lnratio + org.c.in.total.bulk, data = hh_cue)
# cue.cc.ph <- lm(CUE ~ litter.ccratio + soil.ph, data = hh_cue)
# cue.cc.npoc <- lm(CUE ~ litter.ccratio + soil.doc.mass.g, data = hh_cue)
# cue.cc.inorg.n <- lm(CUE ~ litter.ccratio + pct.n.post.ashing.bulk, data = hh_cue)
# cue.cc.orgc <- lm(CUE ~ litter.ccratio + org.c.in.total.bulk, data = hh_cue)

#this model set is redundant, try combining some


#interaction models
``` 


Compare models
*turned off this chunk because these models were mostly redundant
```{ r}
model.set <- list(cue.null, cue.lnratio, cue.litter.c, cue.litter.cc, cue.soil.ph, cue.npoc, cue.inorg.n, cue.org.c, cue.ln.ph, cue.ln.npoc, cue.ln.inorg.n, cue.ln.orgc, cue.cc.ph, cue.cc.npoc, cue.cc.inorg.n, cue.inorg.n, cue.cc.orgc)

modnames <- c("cue.null", "cue.lnratio", "cue.litter.c", "cue.litter.cc", "cue.soil.ph", "cue.npoc", "cue.inorg.n", "cue.org.c", "cue.ln.ph", "cue.ln.npoc", "cue.ln.inorg.n", "cue.ln.orgc", "cue.cc.ph", "cue.cc.npoc", "cue.cc.inorg.n", "cue.inorg.n", "cue.cc.orgc")

(cue.compare <- aictab(cand.set = model.set, modnames = modnames))
``` 


compare just the single models
```{r}
single.model.set <- list(cue.null, cue.lnratio, cue.litter.c, cue.litter.cc, cue.soil.ph, cue.doc, cue.inorg.n, cue.org.c, cue.litter.lignin, cue.org.n)

single.modnames <- c("cue.null", "cue.lnratio", "cue.litter.c", "cue.litter.cc", "cue.soil.ph", "cue.doc", "cue.inorg.n", "cue.org.c", "cue.litter.lignin", "cue.org.n")

(cue.compare <- aictab(cand.set = single.model.set, modnames = single.modnames))

```

Litter carbon carries 42% of the model weight, but the next-best models are still within delta-2 AIC: 

Model selection based on AICc:

                  K    AICc Delta_AICc AICcWt Cum.Wt    LL
cue.litter.c      3 -102.16       0.00   0.42   0.42 54.24
cue.doc           3 -101.04       1.11   0.24   0.67 53.68
cue.inorg.n       3 -100.99       1.16   0.24   0.90 53.66
cue.litter.cc     3  -98.06       4.10   0.05   0.96 52.19
cue.litter.lignin 3  -95.29       6.87   0.01   0.97 50.80
cue.null          2  -94.89       7.26   0.01   0.98 49.53
cue.lnratio       3  -94.82       7.34   0.01   0.99 50.57
cue.soil.ph       3  -92.73       9.43   0.00   1.00 49.54
cue.org.n         3  -91.95      10.20   0.00   1.00 49.14
cue.org.c         3  -90.55      11.61   0.00   1.00 48.44



``` {r}
summary(cue.litter.c)
#A small but significant increase in CUE (with increasing litter C (est = 0.05, p = 0.002).

summary(cue.doc)
#small but significant decline in CUE with increasing DOC in the topsoil (est. = -143.76, p = 0.001), but this relationship seems to be driven by a few outliers.

summary(cue.inorg.n)
#small but significant decrease in CUE with increasing inorganic N (est.=-85.34, p=0.004)
``` 


Plot the data on a graph:
```{r }
(cue.litterc.plot <- ggplot(hh_cue, aes(x = litter.pct.c, y = CUE)) + 
                              geom_point() +
                              geom_smooth(method = "lm", color = " black"))

(doc.cue.plot <- ggplot(hh_cue, aes(x = soil.doc.mass.g, y = CUE)) + 
                              geom_point() +
                              geom_smooth(method = "lm", color = " black"))

(inorg.n.cue.plot <- ggplot(hh_cue, aes(x = pct.n.post.ashing.bulk, y = CUE)) + 
                              geom_point() +
                              geom_smooth(method = "lm", color = " black"))
``` 


Significant variation by litter cc ratio, but a way worse model than model 1 - only explains about half as much variation. Regardless, both have fairly low predictive power. 
```{r }
(cue.litter.cc.plot <- ggplot(hh_cue, aes(x = litter.ccratio, y = CUE)) + 
                              geom_point() +
                              geom_smooth(method = "lm", color = " black"))
``` 


Combine these parameters into a model
```{r }
#cue.lnratio <- lm(CUE ~ lnratio, data = hh_cue)

cue.combined.model <- lm(CUE ~ litter.pct.c + soil.doc.mass.g + pct.n.post.ashing.bulk, data = hh_cue) 

summary(cue.combined.model)
plot(cue.combined.model)
#residual plot shows significant deviations at the edges, is there a better model fit for this?
```


Plot the variation in CUE as function of the 3 main parameters
```{r }
#try a linear regression with litter C as the x axis and soil inorganic N as a factor with 3 levels
summary(hh_cue$pct.n.post.ashing.bulk)
#about a two-fold variation in N 

#make a new dataframe with predicted y values for making the regression
plotting.data <- expand.grid(
  litter.pct.c = seq(min(hh_cue$litter.pct.c), max(hh_cue$litter.pct.c), length.out = 30), 
  soil.doc.mass.g = seq(min(na.omit(hh_cue$soil.doc.mass.g)), max(na.omit(hh_cue$soil.doc.mass.g))),
  pct.n.post.ashing.bulk = c(min(hh_cue$pct.n.post.ashing.bulk), mean(hh_cue$pct.n.post.ashing.bulk), max(hh_cue$pct.n.post.ashing.bulk))
)

#predict new y values
plotting.data$predicted.y <- predict.lm(cue.combined.model, newdata = plotting.data)


#round off the pct.n values
plotting.data$pct.n.post.ashing.bulk <- round(plotting.data$pct.n.post.ashing.bulk, digits = 5)

#make pct n a factor
plotting.data$pct.n.post.ashing.bulk <- as.factor(as.character(plotting.data$pct.n.post.ashing.bulk))

#check the data
summary(plotting.data)

#plot the raw data
(cue.plot <- ggplot(hh_cue, aes(x = litter.pct.c, y = CUE)) +
  geom_point())

#add the regression lines
(cue.plot <- cue.plot + 
    geom_line(data = plotting.data, aes(x=litter.pct.c, y=predicted.y, color=pct.n.post.ashing.bulk), size=1.25))

#add the formatting
(cue.plot <- cue.plot +
    theme_bw()+
    labs(title = "Microbial Carbon-Use Efficiency as a function of litter carbon, \n inorganic nitrogen and dissolved organic carbon \n in the topsoil", 
         x = "Litter Carbon (% of leaf mass)", 
         y = "Microbial CUE", 
         color = "Soil inorganic N \n (% of total \n topsoil (0-5cm) mass)") +
    annotate(geom="text", x = 55.2, y=0.5, label = "CUE = -2.74 + (0.06*litter carbon) + \n (-11.23*soil DOC) + (-83.18*soil inorganic N)"))
``` 


Test the relationship between microbial cue and soil C distribution

First get the normalized diri distribution
```{r }
names(hhout)


levels(hhout$fractionated)
#select out just the data with fraction info
fraction_data <- hhout[hhout$fractionated=="y",]
summary(fraction_data)

#subset out just relevant variables
fraction_data <- select(hhout, c(site, species, depth, fraction.code, org.c.mass.fraction, soil.doc.mass.g))
fraction_data <- fraction_data[complete.cases(fraction_data),]
summary(fraction_data)

#spread the values of fraction mass over fraction code
fraction_wide <- spread(fraction_data, fraction.code, org.c.mass.fraction)
fraction_wide <- fraction_wide[complete.cases(fraction_wide),]

names(fraction_wide)
summary(fraction_wide)
#Make the FD normalized data


fraction.diri <- FD.normalization(fraction_wide[,(4:7)])
fraction.diri


#bind the fraction data to the fraction diri
fraction_wide <- cbind(fraction_wide, fraction.diri)
summary(fraction_wide)

fraction_wide <- fraction_wide[, -c(4:7)]
head(fraction_wide)
```


Join the cue to the diri df
```{r}
fraction_wide_unique <- fraction_wide[unique(fraction_wide$site), ]
summary(fraction_wide_unique)

summary(cue_corrected)
cue_corrected_drop <- cue_corrected[,-1]

#join the cue data to the fraction wide data
fraction_wide_cue <- join(fraction_wide_unique, cue_corrected_drop, type = "left")
#check the join
summary(fraction_wide_cue)
length(fraction_wide_cue$site)
summary(fraction_wide_cue)
``` 


Now do the correlation test
```{r }
orgs <- DR_data(fraction_wide_cue[,4:7])

cue.mod <- DirichReg(orgs ~ CUE*depth, data = fraction_wide_cue)
summary(cue.mod)
``` 

Microbial CUE is not related to soil organic carbon distribution among fractions. 

This suggests that both microbial CUE and soc distribution are responses to litter chemistry and that CUE in itself is not causing variation in soil C distribution. 

What about differences in total C?

First recalculate the tons c/hectare as calculated in the Dirichlet Models script:
```{r }
#first make a cm depth variable
hhout$depth.cm <- revalue(hhout$depth, c("1"="5", "2"="10", "3"="10", "4"="10"))
#summary(hhout$depth)
#summary(hhout$depth.cm)
hhout$depth.cm <- as.numeric(as.character(hhout$depth.cm))
summary(hhout$depth.cm)

#there is a mistake in the bulk density
hhout[hhout$bulk.dens>2, ]

#calculate the tons organic C per hectare
hhout$tons.c.hectare <- hhout$org.c.in.total.bulk * hhout$bulk.dens * hhout$depth.cm
summary(hhout$tons.c.hectare)

#join tons.c.hectare to the hh_cue data
tons.c.for.join <- hhout[, c("site", "species", "depth", "tons.c.hectare")]
hh_cue <- join(hh_cue, tons.c.for.join, type = "left")

summary(hh_cue)
``` 


Next calculate whether CUE is related to tons c per hectare in the top 2 layers
```{r }
cue.c.mass <- lm(tons.c.hectare ~ CUE, data = hh_cue)
summary(cue.c.mass)
#nope
``` 






