---
title: "Dirichlet_distribution"
author: "Becca"
date: "22-5-2019"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Model generation

```{r }
library(tidyverse)
library(plyr)
library(dplyr)
library(ggplot2)
library(FlexDir)
library(tidyr)
```


Make the fraction data into Dirichlet distributions.

load the bayesdf dataframe
```{r }
bayesdf <- read.csv("bayesdf.csv", header=TRUE)
```

First, subset the organic and inorganic fraction masses
```{r}
#subset organic c fraction masses
bayes.sub.org<-select(bayesdf, c(site,species, depth, fraction.code, org.c.fraction.mass))
bso.wide<-spread(bayes.sub.org, fraction.code, org.c.fraction.mass)
names(bso.wide) <- c("site","species", "depth", "org.pom", "org.sa", "org.sc")
summary(bso.wide)

#subset inorganic c fraction masses
bayes.sub.inorg <- select(bayesdf, c(site, species,depth, fraction.code, inorg.c.mass.frac))
bsi.wide<-spread(bayes.sub.inorg, fraction.code, inorg.c.mass.frac)
names(bsi.wide) <- c("site","species", "depth", "inorg.pom", "inorg.sa", "inorg.sc")
summary(bsi.wide)
```


Turn the fraction masses into normalized Dirichlet distributions
```{r }
#install.packages("FlexDir")

bso.wide<-na.omit(bso.wide)
bso.diri <- FD.normalization(bso.wide[,4:6]) 
head(bso.diri)  
bso.w.d<-cbind(bso.wide, bso.diri)

#make sure depth and species are factors
bso.w.d$species <-


bsi.wide<-na.omit(bsi.wide)
bsi.diri <- FD.normalization(bsi.wide[,4:6])
head(bsi.diri)
bsi.w.d<-cbind(bsi.wide, bsi.diri)
```


Paste the data back onto the bayesdf dataframe

First, make the wide version of the bayesdf dataframe.
```{r }
#spread the bulk C data over values of fraction.code
bayesdf.wide <- spread(bayesdf, fraction.code, org.c.bulk.wt)
bayesdf.wide2 <- bayesdf.wide[,1:23]
bayesdf.wide2 <- rename(bayesdf.wide2, replace=c("pom"="org.c.bulk.wt"))
bayesdf.wide2.complete <- na.omit(bayesdf.wide2)
```


Save the dataframe
```{r}
write.csv(bayesdf.wide2.complete, file = "bayesdf.wide2.complete.csv")
```

Read in the data frame
```{r }
bayesdf.wide2.complete <- read.csv("bayesdf.wide2.complete.csv", header=TRUE)

summary(bayesdf.wide2.complete)
``` 


Combine the dirichlet distribution data with the bayesdf.wide2.complete data
```{r}
diri.df <- join(bayesdf.wide2.complete, bsi.w.d[,c(1:3, 7:9)], type = "left")
summary(diri.df)

diri.df <- join(diri.df, bso.w.d[,c(1:3,7:9)], type="left")
summary(diri.df)
```

Model 1, bulk fractions

```{r}
library(DirichletReg)
orgs <- DR_data(bso.w.d[,7:9])
mod1 <- DirichReg(orgs ~ species*depth, data = bso.w.d[, c(1:3, 7:9)])
mod2 <- DirichReg(orgs ~ species + depth, data = bso.w.d[,c(1:3, 7:9)])
summary(mod1)
summary(mod2)

AIC(mod1)
AIC(mod2)
#picea is quite different from all other species in terms of organic C distribution. 
```


Plot the organic C distributions by species at depth 1.
```{r}

bso.w.d1<-subset(bso.w.d, depth==1)
orgs1 <- DR_data(bso.w.d1[,7:9])

rangens <- function(sp.name, depth="1", n= 1000){
            alphas <-matrix(predict(mod1, alpha=TRUE)$alpha[bso.w.d$species==sp.name &
                            bso.w.d$depth==depth,], ncol=3)[1,]
            rands <- rdirichlet(n, alphas)
            return(rands)
}

#apply the random generation to each level of the 'species' variable
a <- lapply(levels(bso.w.d1$species), rangens)
names(a) <- levels(bso.w.d1$species)
summary(a)
###to apply to the variable 'fraxinus', specify length as 5, do the following:
#rangens("fraxinus", n = 5)
###
library(car)
library(RColorBrewer)
mycols<-brewer.pal(n=6, name="Accent")
mycols2<-adjustcolor(mycols, alpha.f=0.1)
mycols3<-brewer.pal(n=6, name="Dark2")


plot.new()
plot(orgs1,  reset_par = FALSE, type="n")
lines(dataEllipse(toTernary(DR_data(a[[1]])), draw=FALSE)$`0.95`,col=mycols3[[1]])
lines(dataEllipse(toTernary(DR_data(a[[2]])), draw=FALSE)$`0.95`,col=mycols3[[2]])
lines(dataEllipse(toTernary(DR_data(a[[3]])), draw=FALSE)$`0.95`,col=mycols3[[3]])
lines(dataEllipse(toTernary(DR_data(a[[4]])), draw=FALSE)$`0.95`,col=mycols3[[4]])
lines(dataEllipse(toTernary(DR_data(a[[5]])), draw=FALSE)$`0.95`,col=mycols3[[5]])
lines(dataEllipse(toTernary(DR_data(a[[6]])), draw=FALSE)$`0.95`,col=mycols3[[6]])

#1=Acer, 2=fagus, 3=fraxinus, 4=picea, 5=populus, 6=quercus
```


Do the same for depth 2
```{r}

bso.w.d2<-subset(bso.w.d, depth==2)
orgs2 <- DR_data(bso.w.d1[,7:9])

rangens2 <- function(sp.name, depth="2", n= 1000){
            alphas <-matrix(predict(mod1, alpha=TRUE)$alpha[bso.w.d$species==sp.name &
                            bso.w.d$depth==depth,], ncol=3)[1,]
            rands <- rdirichlet(n, alphas)
            return(rands)
}

#apply the random generation to each level of the 'species' variable
b <- lapply(levels(bso.w.d2$species), rangens2)
names(b) <- levels(bso.w.d2$species)
summary(b)
###to apply to the variable 'fraxinus', specify length as 5, do the following:
#rangens("fraxinus", n = 5)
###

plot.new()
plot(orgs2,  reset_par = FALSE, type="n")
lines(dataEllipse(toTernary(DR_data(b[[1]])), draw=FALSE)$`0.95`,col=mycols3[[1]])
lines(dataEllipse(toTernary(DR_data(b[[2]])), draw=FALSE)$`0.95`,col=mycols3[[2]])
lines(dataEllipse(toTernary(DR_data(b[[3]])), draw=FALSE)$`0.95`,col=mycols3[[3]])
lines(dataEllipse(toTernary(DR_data(b[[4]])), draw=FALSE)$`0.95`,col=mycols3[[4]])
lines(dataEllipse(toTernary(DR_data(b[[5]])), draw=FALSE)$`0.95`,col=mycols3[[5]])
lines(dataEllipse(toTernary(DR_data(b[[6]])), draw=FALSE)$`0.95`,col=mycols3[[6]])

#1=Acer, 2=fagus, 3=fraxinus, 4=picea, 5=populus, 6=quercus

```


make a stacked plot
```{r}
#install.packages("wesanderson")
library(wesanderson)
fill=wes_palette("FantasticFox1", 3)

library(tidyr)
diri.df.long <- diri.df%>%
  select(species, depth, org.pom, org.sa, org.sc)%>%
  group_by(species, depth)%>%
  slice(1)
diri.df.long <- diri.df.long%>%gather(key="fraction", value="weight", -species, -depth)  

diri.df.long$fraction <- as.factor(diri.df.long$fraction)  
summary(diri.df.long)  

sp.orgc<-diri.df.long %>%
  group_by(species) %>%
  summarize(total_c = sum(weight))
sp.orgc

c.plot.species <- ggplot(diri.df.long, aes(x = species, y = weight, fill = fraction)) + 
  geom_col() +
  theme(legend.title = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Dominant Species", y = "Proportion of dry-weight OC contributed by solid fraction")

c.plot.species + 
  facet_wrap(~depth) +
  scale_fill_manual(values=fill)
```


Plot organic C weights (sum of fractions)
```{r}
orgc.totwt <- ggplot(diri.df, aes(x=species, y=totwt)) +
  geom_point() +
  geom_boxplot(alpha=0.3) +
  theme(legend.title = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Dominant Species", y = "Total organic C (sum of fractions)") 

orgc.totwt +
  facet_wrap(~depth) +
  geom_hline(yintercept=0.482, linetype="dashed", color = "red")
```


Plot organic C weights (total bulk)
```{r }
orgc.bulkwt <- ggplot(diri.df, aes(x=species, y=org.c.start.mass)) +
  geom_point() +
  geom_boxplot(alpha=0.3) +
  theme(legend.title = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Dominant Species", y = "Total organic C (g) (bulk soil)") 

orgc.bulkwt +
  facet_wrap(~depth) +
  #mean bulk mass
  geom_hline(yintercept=0.446, linetype="dashed", color = "red", size=1) + 
  #mean sum of fractions mass
  geom_hline(yintercept=0.482, linetype="dashed", color="black")
```


Test the relationship between species, depth, fractionation status, and organic c total weight.
```{r }
#data
fracwts <- select(diri.df, site, depth, species, totwt)
fracwts$frac <- 1


bulkwts <- select(diri.df, site, depth, species, org.c.start.mass)
bulkwts$frac <- 2
bulkwts$totwt <- bulkwts$org.c.start.mass
bulkwts <- select(bulkwts, site, depth, species, totwt, frac)

orgwts <- rbind(fracwts, bulkwts)
summary(orgwts)  
orgwts$frac <- as.factor(orgwts$frac)  

##multimodel comparison##

#install.packages("AICcmodavg")
library(AICcmodavg)

Cand.set <- list(
  lm(totwt ~ 1, data=orgwts),
  lm(totwt ~ species, data=orgwts),
  lm(totwt ~ depth, data=orgwts),
  lm(totwt ~ frac, data=orgwts),
  lm(totwt ~ species + depth + frac, data = orgwts),
  lm(totwt ~ species*depth + frac, data = orgwts)
)

aictab(cand.set = Cand.set, second.ord = FALSE)
```


Model 6 carries 97% of the cumulative AIC weight. Models 2 and 4 do worse than random.

Model 6
```{r }
totc6<-lm(totwt ~ species*depth + frac, data = orgwts)
summary(totc6)

library(lme4)
totc6.lme <- lmer(totwt ~ species*depth + frac + (1|site), data = orgwts)
summary(totc6.lme)
emmeans(totc6.lme, list(pairwise ~ species*depth), adjust= "tukey")

library(emmeans)
orgwts_bulk <- orgwts[orgwts$frac==2,]
totc6.lme.bulk <- lmer(totwt ~ species*depth + (1|site), data = orgwts_bulk)
emmeans(totc6.lme.bulk, list(pairwise ~ species*depth), adjust="tukey")


orgwts_frac <- orgwts[orgwts$frac==1,]
totc6.lme.frac <- lmer(totwt ~ species*depth + (1|site), data = orgwts_frac)
emmeans(totc6.lme.frac, list(pairwise ~ species*depth), adjust="tukey")

orgwts_depth1 <- orgwts[orgwts$depth==1,] # don't do this! P hack alert!
orgwts_depth2 <- orgwts[orgwts$depth==2,]

totc6.lme1 <- lmer(totwt ~ species + frac + (1|site), data = orgwts_depth1)
totc6.lme2 <- lmer(totwt ~ species + frac + (1|site), data = orgwts_depth2)


#posthoc means comparison
emmeans(totc6.lme1, list(pairwise ~ species), adjust = "tukey")
emmeans(totc6.lme2, list(pairwise ~ species), adjust = "tukey")

```



```{ r}
orgwts$interaction <- interaction(orgwts$species, orgwts$depth)
multComp(totc6_mle4, "interaction")

orgwts_depth1 <- orgwts[orgwts$depth==1,] # don't do this! P hack alert!
orgwts_depth2 <- orgwts[orgwts$depth==2,]

wt.mod1 <- lm(totwt ~ species + frac , data = orgwts_depth1)
summary(wt.mod1)
multComp(wt.mod1, "species")

wt.mod2 <- lm(totwt ~ species + frac, data=orgwts_depth2)
multComp(wt.mod2, "species")
```



Hacky QRPs ahoy:
```{r}
bso.w.d$pom_sc <- bso.w.d[,7] / bso.w.d[,9]

pomsc.rag<-lm(pom_sc ~ species * depth, data=bso.w.d)
aov.pomsc <- aov(pomsc.rag)
TukeyHSD(aov.pomsc)
  
#Anova(pomsc.rag)
#multComp(lm(pom_sc ~ species + depth, bso.w.d), "species") # not realistic because no interactions
```


Test relationship between litter lignin:N ratio and pom_sc ratio
```{r }
litter <- read.csv("litter_cn.csv", header=TRUE)
summary(litter)
#litter$lnratio <- as.numeric(litter$lnratio)

#litter$lnratio <- (litter$lignin.mg.g/100)/litter$pct.n
#summary(litter)

#write.csv(litter, file="litter_cn.csv")

#add litter ln ratio to the bso.w.d dataframe
bso.w.d.sub<-bso.w.d[,c(1:3, 7:10)]

#add mean values of litter ln ratio 
bso.w.d.sub$lnratio <- bso.w.d.sub$species
bso.w.d.sub$lnratio <- mapvalues(bso.w.d.sub$lnratio, from = c("fraxinus", "quercus", "acer", "populus", "fagus", "picea"), to=c("0.531", "2.030", "1.154", "1.149", "2.444", "2.144"))
bso.w.d.sub$lnratio <- as.numeric(as.character(bso.w.d.sub$lnratio))
summary(bso.w.d.sub)
```


Run regression analysis again using ln ratio instead of species
```{r }
pom_sc_lm<-lm(pom_sc ~ lnratio + depth, data=bso.w.d.sub)
pom_sc_lm2<-lm(pom_sc ~ lnratio*depth, data=bso.w.d.sub)

AIC(pom_sc_lm)
AIC(pom_sc_lm2)

summary(pom_sc_lm)
```

ggplot(data = df, aes(x = x, y = y, label=y)) +
  stat_smooth_func(geom="text",method="lm",hjust=0,parse=TRUE) +
  geom_smooth(method="lm",se=FALSE) +
  geom_point() + facet_wrap(~class)
  
Plot the lm
```{r}
#library(devtools)
#source_gist("524eade46135f6348140")

pom_sc_plot <- ggplot(bso.w.d.sub, aes(x=lnratio, y=pom_sc)) +
  geom_point() +
  geom_jitter(position = position_jitter(width = .15)) +
  #stat_smooth_func(geom="text", color="black", method="lm", hjust=0, vjust=-1.3, parse=TRUE) +
  geom_smooth(method="lm", color="black") +
  theme(legend.title = element_blank(), axis.text.x = element_text()) +
  labs(x = "Litter lignin:N ratio", y = "Soil POM:SC ratio") 
pom_sc_plot + facet_wrap(~depth)


#pom_sc2 <- ggplot(bso.w.d.sub, aes())
```


Get difference in start and final soil c
```{r }
#get the values
diri.df$pct.diff <- ((diri.df$totwt - diri.df$org.c.start.mass)/diri.df$totwt)*100

#plot the difference
pct.diff <- ggplot(diri.df, aes(x=species, y=pct.diff)) +
  geom_point() +
  geom_boxplot(alpha=0.3) +
  theme(legend.title = element_blank(), axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "Dominant Species", y = "Difference between sum-of-fractions and bulk org C (%)") 

pct.diff +
  facet_wrap(~depth) 

pctdiff.lm<-lm(pct.diff ~ species*depth, data=diri.df)
summary(pctdiff.lm)
```

