---
title: "Soil_C_Distribution_for_paper"
author: "Becca"
date: "3/12/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro


The purpose of this file is to remake the soil C distribution plots and statistics using the master data frame rather than the individual frames used in Dirichlet_distributions.

The file will first 

## 1: Setup 

Load packages
```{r }
library(tidyverse)
library(tidyr)
library(plyr)
library(dplyr)
library(ggplot2)
library(FlexDir)
library(DirichletReg)
library(AICcmodavg)
library(ape)
library(mvtnorm)
```


Load data
```{r }
hhout <- read.csv("hhout_master_01May2020.csv", header=TRUE)
hhout$depth <- as.factor(hhout$depth)
bayesdf <- read.csv("bayesdf.csv", header=TRUE)
summary(hhout)
summary(bayesdf)

#this comparison isn't working, but if you look at the summaries they're equal.
#all.equal(hhout[hhout$fractionated=="y", ]$org.c.fraction.mass, bayesdf$org.c.fraction.mass) 
```


## 2: Building intermediate dataframes - skip ahead to section 3

Subset out the organic and inorganic fraction masses
```{ r }
#subset organic c fraction masses
sub.org<-select(hhout, c(site, species, depth, fraction.code, org.c.fraction.mass))
summary(sub.org)
sub.org <- sub.org[complete.cases(sub.org), ]
summary(sub.org)

sub.org.wide<-spread(sub.org, fraction.code, org.c.fraction.mass)
names(sub.org.wide) <- c("site","species", "depth", "org.pom", "org.sa", "org.sc")
summary(sub.org.wide)

#subset inorganic c fraction masses
sub.inorg <- select(hhout, c(site, species,depth, fraction.code, inorg.c.mass.frac))
summary(sub.inorg)
sub.inorg <- sub.inorg[complete.cases(sub.inorg), ]
summary(sub.inorg)

sub.inorg.wide <- spread(sub.inorg, fraction.code, inorg.c.mass.frac)
names(sub.inorg.wide) <- c("site","species", "depth", "inorg.pom", "inorg.sa", "inorg.sc")
summary(sub.inorg.wide)
```


Turn the masses into normalized Dirichlet distributions
```{ r }
#organic
sub.org.wide <- na.omit(sub.org.wide)
org.diri <- FD.normalization(sub.org.wide[,4:6]) 
head(org.diri)  
sub.org.wide <- cbind(sub.org.wide, org.diri)


#inorganic
sub.inorg.wide <- na.omit(sub.inorg.wide)
inorg.diri <- FD.normalization(sub.inorg.wide[,4:6])
head(inorg.diri)
sub.inorg.wide <- cbind(sub.inorg.wide, inorg.diri)
```


Subset the master data frame
```{ r}
#first get just the rows from the fraction data
hhout_sub <- hhout[hhout$fractionated=="y", ] 

#now pull out just the fraction.code, org.c.bulk.wt, site, species, depth, pct.mass.loss, soil.start.wt, weight.g, orgc.weight, totwt, org.c.start.mass, org.c.fraction.mass, inorg.c.mass.frac, inorg.c.mass.bulk
hhout_sub <- hhout_sub %>% select("fraction.code", "org.c.start.mass.bulk", "site", "species", "depth", "pct.mass.loss.fraction", "pct.mass.loss.bulk", "soil.mass.pre.fract", "fraction.mass", "org.c.mass.fraction", "sum.org.c.mass.fraction", "inorg.c.mass.frac", "inorg.c.mass.bulk")


#spread the bulk C, pct.n.pre.ashing, pct.n.post.ashing, pct.mass.loss, c.in.org.fract, org.c.in.t.fract, org.n.in.t.frac, weight.g, orgc.weight, pct.c.cont.by.fraction, org.c.fraction.mass, inorg.c.mass.frac variables over values of fraction.code
hhout_sub_wide <- spread(hhout_sub, fraction.code, org.c.start.mass.bulk)
summary(hhout_sub_wide)

#rename one of the fractions with the org.c.bulk.wt
hhout_sub_wide <- rename(hhout_sub_wide, replace = c("pom" = "org.c.start.mass.bulk"))
summary(hhout_sub_wide)

#omit the last two columns
hhout_sub_wide_drop <- hhout_sub_wide[ , -c(13, 14)]
summary(hhout_sub_wide_drop)

#drop the nas
hhout_sub_wide_complete <- na.omit(hhout_sub_wide_drop)
summary(hhout_sub_wide_complete)
```


Paste the data back into the dirichlet master dataframe
```{ r }
diri_df <- join(hhout_sub_wide_complete, sub.org.wide[ ,c(1:3, 7:9)], type="left")
diri_df <- join(diri_df, sub.inorg.wide[ ,c(1:3, 7:9)], type="left")

summary(diri_df)
```


Write the new dataframes so we don't have to keep redoing the above steps.
```{ r }
write.csv(diri_df, file = "diri_df.csv", row.names = FALSE)
write.csv(hhout_sub_wide_complete, file = "hhout_sub_wide_complete.csv", row.names = FALSE)
write.csv(hhout_sub_wide, file = "hhout_sub_wide.csv", row.names = FALSE)
write.csv(sub.org.wide, file = "sub.org.wide.csv", row.names = FALSE)
```


The soil npoc data is on a different scale from the other variables (mg.L vs g/g). Redo to make the scale the same as the other variables (g/dry mass)
```{ r }
# (mg/L * volume of extract (L))/dw soil in extract
hhout$npoc.mg.g <- (hhout$soil.npoc.mg.L * 0.04) / 5
summary(hhout$npoc.mg.g)

#change mg.g to g to match other soc numbers
hhout$soil.doc.mass.g <- hhout$npoc.mg.g/100
summary(hhout$soil.doc.mass.g)

# check whether this correlates with anything else
soil.doc.lm <- lm(npoc.mg.g ~ soil.ph + litter.pct.c + sum.org.c.mass.fraction, data=hhout)
summary(soil.doc.lm)

write.csv(hhout, file = "hhout_master_01May2020.csv", row.names = FALSE)
```

So DOC increases along with soil pH, and decreases with sum organic c mass fraction (makes sense, since more DOC would mean less C in the solid fractions and vice-versa). 


## 3: Quantify the distribution of soil organic carbon by litter chemistry and depth

Build the dataframe
```{r}
#subset the organic fraction values for the dirichlet model
sub.org.wide <- read.csv("sub.org.wide.csv", header = TRUE)
sub.org.wide$depth = as.factor(sub.org.wide$depth)

npoc.for.join <- hhout[,c("site", "species", "depth","soil.ph", "soil.doc.mass.g")]
sub.org.wide <- join(sub.org.wide, npoc.for.join, type="left")

litter.for.join <- hhout[,c("site", "species", "depth", "litter.pct.c", "litter.pct.n", "litter.lignin.mg.g")]
sub.org.wide <- join(sub.org.wide, litter.for.join, type="left")

sub.org.wide %>% distinct() -> sub.org.wide

summary(sub.org.wide)

orgs <- DR_data(sub.org.wide[,c(7:9, 11)])
```


Create some models to compare influence of different litter traits on C distribution. Use depth as an interactive factor because depth is very likely to influence the effect of some of these parameters on C distribution. 
```{r }
#make a model with species and depth as predictors
c.mod1 <- DirichReg(orgs ~ species*depth, data = sub.org.wide, model = "alternative")

#use litter chemistry instead of species as a predictor
c.mod2 <- DirichReg(orgs ~ litter.pct.n*depth, data = sub.org.wide, model = "alternative")
c.mod3 <- DirichReg(orgs ~ litter.pct.c*depth, data = sub.org.wide, model = "alternative")
c.mod4 <- DirichReg(orgs ~ litter.lignin.mg.g*depth, data = sub.org.wide, model = "alternative")

#combine lignin and n into lnratio
sub.org.wide$lnratio <- ((sub.org.wide$litter.lignin.mg.g/10)/sub.org.wide$litter.pct.n)
summary(sub.org.wide$lnratio)

c.mod5 <- DirichReg(orgs ~ lnratio*depth, data = sub.org.wide, model = "alternative")

#see if soil ph is related to c distribution
c.mod6 <- DirichReg(orgs ~ soil.ph*depth, data = sub.org.wide, model = "alternative")

#run the model summaries
#summary(mod1)
summary(c.mod1)
summary(c.mod2)
summary(c.mod3)
summary(c.mod4)
summary(c.mod5)
summary(c.mod6)

#compare AICs
#AIC(mod1)
AIC(c.mod1)
AIC(c.mod2)
AIC(c.mod3)
AIC(c.mod4)
AIC(c.mod5)
AIC(c.mod6)
```

The best moels (all within 1 delta-AIC of each other) are 5, 4, and 2 (lnratio, lignin, and nitrogen). I will move forward with model 5 (lnratio).

Call:
DirichReg(formula = orgs ~ lnratio * depth, data = sub.org.wide, model = "alternative")

Standardized Residuals:
                     Min       1Q   Median       3Q     Max
org.pom.1        -3.0530  -0.4907  -0.2485   0.8173  5.1884
org.sa.1         -2.0787  -0.6160  -0.0119   0.7225  3.0999
org.sc.1         -3.3082  -0.6740   0.1294   0.6713  2.5103
soil.doc.mass.g  -0.7478  -0.5649  -0.4818  -0.3522  0.1802

MEAN MODELS:
------------------------------------------------------------------
Coefficients for variable no. 1: org.pom.1
- variable omitted (reference category) -
------------------------------------------------------------------
Coefficients for variable no. 2: org.sa.1
               Estimate Std. Error z value Pr(>|z|)   
(Intercept)     0.10086    0.20949   0.481  0.63021   
lnratio        -0.03266    0.01198  -2.725  0.00643 **
depth2         -0.08988    0.34235  -0.263  0.79291   
lnratio:depth2  0.01830    0.01940   0.943  0.34557   
------------------------------------------------------------------
Coefficients for variable no. 3: org.sc.1
                Estimate Std. Error z value Pr(>|z|)    
(Intercept)     1.398550   0.162282   8.618  < 2e-16 ***
lnratio        -0.023472   0.009053  -2.593  0.00952 ** 
depth2          0.536851   0.258437   2.077  0.03777 *  
lnratio:depth2  0.003254   0.014431   0.225  0.82160    
------------------------------------------------------------------
Coefficients for variable no. 4: soil.doc.mass.g
                Estimate Std. Error z value Pr(>|z|)    
(Intercept)    -2.407391   0.419137  -5.744 9.26e-09 ***
lnratio        -0.021445   0.023879  -0.898    0.369    
depth2          0.546588   0.599973   0.911    0.362    
lnratio:depth2  0.006925   0.034186   0.203    0.839    
------------------------------------------------------------------

PRECISION MODEL:
------------------------------------------------------------------
            Estimate Std. Error z value Pr(>|z|)    
(Intercept)   3.9591     0.1088   36.38   <2e-16 ***
------------------------------------------------------------------
Significance codes: 0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Log-likelihood: 385.8 on 13 df (90 BFGS + 2 NR Iterations)
AIC: -745.7, BIC: -718.5
Number of Observations: 60
Links: Logit (Means) and Log (Precision)
Parametrization: alternative

## 4: Visualize the results of the Dirichlet distribution 

For visualization purposes, predict the values of each fraction across the observed lnratio values, then plot the moel fit against the observed data.

First prepare the prediction data
```{r }
#turn the sub.org.wide data from wide to long
sub.org.long <- sub.org.wide%>%pivot_longer(
  cols = c(soil.doc.mass.g, org.pom.1, org.sa.1, org.sc.1), 
  names_to = "fraction.code", 
  values_to = "proportion")

#make the factors back into factors
sub.org.long$fraction.code <- as.factor(sub.org.long$fraction.code)
sub.org.long$depth <- as.factor(sub.org.long$depth)
summary(sub.org.long)


#make depth a number for this step 
sub.org.long$depth <- as.numeric(sub.org.long$depth)

#first predict the values from the dirichlet model
dirichlet.plot.data <- expand.grid( 
  lnratio = seq(min(sub.org.long$lnratio), max(sub.org.long$lnratio), length.out = 100),
  depth = c(min(sub.org.long$depth), max(sub.org.long$depth))) 
summary(dirichlet.plot.data)


#make depth a factor again, clean up the data frame
dirichlet.plot.data$depth <- as.factor(dirichlet.plot.data$depth)
dirichlet.plot.data %>% distinct() -> dirichlet.plot.data
summary(dirichlet.plot.data)
length(dirichlet.plot.data$lnratio)

#predict the values of the different variables under the dirichlet model
predicted.data <- predict.DirichletRegModel(c.mod5, newdata = dirichlet.plot.data, type = "link", se = TRUE)
#V1 = POM
#V2 = SA
#V3 = SC
#V4 = DOC

#make the matrix into a data frame
predicted.data <- as.data.frame(predicted.data)
summary(predicted.data)

#change names to fraction code
predicted.data <- predicted.data%>%dplyr::rename(
  pom = V1, 
  sa = V2,
  sc = V3,
  doc = V4)
summary(predicted.data)
is.data.frame(predicted.data)

#combine the predicted.data and the dirichlet plot data using cbind
dirichlet.plot.data <- cbind(dirichlet.plot.data, predicted.data)
summary(dirichlet.plot.data)

#turn the predicted values from wide to long
dirichlet.plot.data <- dirichlet.plot.data%>%pivot_longer(
  cols = c(pom, sa, sc, doc), 
  names_to = "fraction", 
  values_to = "proportion")

dirichlet.plot.data$fraction <- as.factor(dirichlet.plot.data$fraction)
summary(dirichlet.plot.data)
```


Plot the raw data and model fits
```{r }
#subset the hhout data for plotting
hhout.sub <- hhout[hhout$fractionated=="y", ]

#geom_line(data = plotting.data, aes(x=litter.pct.c, y=predicted.y, color=pct.n.post.ashing.bulk), size=1.25))
jitter <- position_jitter(width = 0.5)

(dirichlet.plot <- ggplot(sub.org.long, aes(x=lnratio, y=proportion)) +
                        geom_jitter(aes(color = fraction.code)) +
                        geom_smooth(data = dirichlet.plot.data, aes(x=lnratio, y=proportion, color = fraction))+
                        facet_wrap(~depth) +
                        theme_bw())
```

*WORKING* Plot same data (for depth 1) including bootstrap confidence regions in 'base' graphics.

```{r}

model.mean <- unlist(coef(c.mod5))
model.sig <- vcov(c.mod5)

sims <- rmvnorm(1, mean = model.mean, sigma = diag(x = 0,nrow = 13, ncol = 13))



# depth1 only
pred.matrix <- expand.grid(intercept = 1, lnratio = seq(0,25,length.out = 100),
                            depth2 = 0, interaction = 0)

pred.values <- cbind(as.matrix(pred.matrix) %*% sims[1:4],
                     as.matrix(pred.matrix) %*% sims[5:8],
                     as.matrix(pred.matrix) %*% sims[9:12])


# start plot with pom
plot(orgs[,1] ~ sub.org.wide$lnratio,
     ylim = c(0,1), xlim = c(5,30),
     ylab = "Proportion",
     xlab = "lignin-N ratio")

lines(dirichlet.plot.data$lnratio[dirichlet.plot.data$fraction=="pom"&dirichlet.plot.data$depth ==1], dirichlet.plot.data$proportion[dirichlet.plot.data$fraction=="pom"&dirichlet.plot.data$depth ==1],
      lwd = 2)




```

## 5: Plot the organic C distributions (solid fractions) by species at depths 1 and 2 using ternary plots

I have turned all of these chunks off because I don't think that they're useful or interesting, plus they do not show the whole distribution. 

Skip to section 6

First depth 1
```{ r }
org.wide.1<-subset(sub.org.wide, depth==1)

orgs1 <- DR_data(org.wide.1[,7:9])

rangens <- function(sp.name, depth="1", n= 1000){
            alphas <-matrix(predict(mod2, alpha=TRUE)$alpha[org.wide.1$species==sp.name &
                            org.wide.1$depth==depth,], ncol=3)[1,]
            rands <- rdirichlet(n, alphas)
            return(rands)
}

#apply the random generation to each level of the 'species' variable
a <- lapply(levels(org.wide.1$species), rangens)
names(a) <- levels(org.wide.1$species)
summary(a)
###to apply to the variable 'fraxinus', specify length as 5, do the following:
#rangens("fraxinus", n = 5)
###
library(car)
library(RColorBrewer)
mycols<-brewer.pal(n=6, name="Accent")
mycols2<-adjustcolor(mycols, alpha.f=0.1)
mycols3<-brewer.pal(n=6, name="Dark2")


plot.new()
plot(orgs1,  reset_par = FALSE, type="n")
lines(dataEllipse(toTernary(DR_data(a[[1]])), draw=FALSE)$`0.95`,col=mycols3[[1]])
lines(dataEllipse(toTernary(DR_data(a[[2]])), draw=FALSE)$`0.95`,col=mycols3[[2]])
lines(dataEllipse(toTernary(DR_data(a[[3]])), draw=FALSE)$`0.95`,col=mycols3[[3]])
lines(dataEllipse(toTernary(DR_data(a[[4]])), draw=FALSE)$`0.95`,col=mycols3[[4]])
lines(dataEllipse(toTernary(DR_data(a[[5]])), draw=FALSE)$`0.95`,col=mycols3[[5]])
lines(dataEllipse(toTernary(DR_data(a[[6]])), draw=FALSE)$`0.95`,col=mycols3[[6]])
#1=Acer, 2=fagus, 3=fraxinus, 4=picea, 5=populus, 6=quercus
```


Do the same for depth 2
```{ r }
org.wide.2<-subset(sub.org.wide, depth==2)

orgs2 <- DR_data(org.wide.2[,7:9])

rangens2 <- function(sp.name, depth="2", n= 1000){
            alphas <-matrix(predict(mod2, alpha=TRUE)$alpha[org.wide.2$species==sp.name &
                            org.wide.2$depth==depth,], ncol=3)[1,]
            rands <- rdirichlet(n, alphas)
            return(rands)
}

#apply the random generation to each level of the 'species' variable
b <- lapply(levels(org.wide.2$species), rangens2)
names(b) <- levels(org.wide.2$species)
summary(b)
df.b <- as.data.frame(b)
summary(df.b)
###to apply to the variable 'fraxinus', specify length as 5, do the following:
#rangens("fraxinus", n = 5)
###

plot.new()
plot(orgs2,  reset_par = FALSE, type="n")
lines(dataEllipse(toTernary(DR_data(b[[1]])), draw=FALSE)$`0.95`,col=mycols3[[1]])
lines(dataEllipse(toTernary(DR_data(b[[2]])), draw=FALSE)$`0.95`,col=mycols3[[2]])
lines(dataEllipse(toTernary(DR_data(b[[3]])), draw=FALSE)$`0.95`,col=mycols3[[3]])
lines(dataEllipse(toTernary(DR_data(b[[4]])), draw=FALSE)$`0.95`,col=mycols3[[4]])
lines(dataEllipse(toTernary(DR_data(b[[5]])), draw=FALSE)$`0.95`,col=mycols3[[5]])
lines(dataEllipse(toTernary(DR_data(b[[6]])), draw=FALSE)$`0.95`,col=mycols3[[6]])
#1=Acer, 2=fagus, 3=fraxinus, 4=picea, 5=populus, 6=quercus
```


Try redoing these ternary plots with ggtern
```{ r }
library(ggtern)

#depth 1
ggtern(data=org.wide.1, aes(org.pom.1, org.sc.1, org.sa.1)) + theme_showarrows() + geom_point(aes(color=species)) + theme_dark()

#depth 2
ggtern(data=org.wide.2, aes(org.pom.1, org.sc.1, org.sa.1)) + theme_showarrows() + geom_point(aes(color=species)) 
```

Okay, I'm done with this, I don't think the ternary plots are interesting or helpful.

## 6: Visualize litter chemistry parameters by species (figure 4 in paper)

First prepare the data frame
```{r }
#read in the hhout_sub_wide_complete dataframe
hhout_sub_wide_complete <- read.csv("hhout_sub_wide_complete.csv", header = TRUE)
hhout_sub_wide_complete$depth <- as.factor(hhout_sub_wide_complete)
summary(hhout_sub_wide_complete)

hhout_sub_wide_complete %>% distinct() -> hhout_sub_wide_complete

#check names of the columns
names(sub.org.wide)
names(hhout_sub_wide_complete)

#subset litter chemistry and npoc data from the complete hhout data frame
litter.dat <- hhout[hhout$fractionated=="y", c(1,2,14,42:46,50)] 
summary(litter.dat)


#join the litter and npoc data to the diri data
sub.org.litter <- join(sub.org.wide, litter.dat, type = "full")
sub.org.litter %>% distinct() -> sub.org.litter

#make the lignin:n ratio variable
sub.org.litter$lnratio <- (sub.org.litter$litter.lignin.mg.g/10)/sub.org.litter$litter.pct.n
summary(sub.org.litter$lnratio)

#make the soil pom:sc ratio
sub.org.litter$pom.sc.ratio <- sub.org.litter$org.pom/sub.org.litter$org.sc
summary(sub.org.litter$pom.sc.ratio)


#add the litter hwec.mg.g to the dataframe from the hh_cue dataframe
hh_cue <- read.csv("hh_cue.csv", header = TRUE)
sub.for.join <- hh_cue[,c(1,14,55)]
sub.org.litter <- join(sub.org.litter, sub.for.join, type = "left")

sub.org.litter %>% distinct() -> sub.org.litter

summary(sub.org.litter)

```


Plot lnratio, litter c, litter n, and hwe-c by species
```{r }
sub.org.litter$species <- with(sub.org.litter, reorder(species, lnratio, mean))
  
(ln.species <- ggplot(sub.org.litter, aes(x=species, y=lnratio)) +
   geom_point() +
   theme_bw())

#sub.org.litter$species <- with(sub.org.litter, reorder(species, litter.hwec.mg.g, mean))

(hwec.species <- ggplot(sub.org.litter, aes(x=species, y=litter.hwec.mg.g)) +
    geom_point() +
    theme_bw())

#sub.org.litter$species <- with(sub.org.litter, reorder(species, litter.pct.c, mean))

(pctc.species <- ggplot(sub.org.litter, aes(x=species, y=litter.pct.c)) +
    geom_point() +
    theme_bw())

#sub.org.litter$species <- with(sub.org.litter, reorder(species, litter.pct.n, mean))

(pctn.species <- ggplot(sub.org.litter, aes(x=species, y=litter.pct.n)) +
    geom_point() +
    theme_bw())
```


Put the above plots together in a plot
```{r }
library(gridExtra)
grid.arrange(ln.species, pctc.species, pctn.species, hwec.species, ncol = 2)

```


check for autocorrelation among independent variables - I have turned off this block because the ggpairs() function can cause the script to hang.  
```{ r}
#install.packages("GGally")
library(GGally)

#correlation matrix
ggpairs(sub.org.litter[,c(2, 3, 7:14)])

#values that shouldn't go into the same model: 
#  litter ln ratio and lignin, litter pct c, litter pct n, litter lignin mg, litter cell mg, litter hwe c ppm, 
#  litter pct c and lignin, hwec, and lnratio
#  litter hwec ppm and litter ln ratio (corr -0.98)
```


## 7: Look at variation in the distribution of soil C fractions using distance matrices

This is an alternate method to the dirichlet distribution for calculating differences in soil fractions according to litter chemistry. 

```{ r }
library(vegan)

#transform compositional data
##first put the non-normalized organic masses back into the dataframe
pcoa_df <- join(hhout_sub_wide_complete, sub.org.wide[ ,c(1:6,10)], type="left")
summary(pcoa_df)
pcoa_df$depth = as.factor(pcoa_df$depth)

#add the soil hwec mg g 
hwec.for.join <- litter.dat[,c("site", "species", "depth", "soil.doc.mass.g")]
pcoa_df <- join(pcoa_df, npoc.for.join, type="left")
summary(pcoa_df)

pcoa_sub <- pcoa_df[ ,c(13:16)]
summary(pcoa_sub)
  
##next get the distance matrix using vegdist
#NEED TO ADD NPOC TO THIS
soil.D<-vegdist(pcoa_sub, "bray", na.rm = TRUE)

#do the pcoa
soilpcoa <- pcoa(soil.D, correction = "cailliez")
#soilpcoa

#plot the pcoa
biplot(soilpcoa)
```

The PCoA suffered from the horseshoe effect.

Try an NMDS with envfit()
```{ r }
#check the stress vs. dimension plot
NMDS.scree <- function(x) { #where x is the name of the data frame variable
  plot(rep(1, 10), replicate(10, metaMDS(x, autotransform = F, k = 1)$stress), xlim = c(1, 10),ylim = c(0, 0.30), xlab = "# of Dimensions", ylab = "Stress", main = "NMDS stress plot")
  for (i in 1:10) {
    points(rep(i + 1,10),replicate(10, metaMDS(x, autotransform = F, k = i + 1)$stress))
  }
}

NMDS.scree(soil.D)
#2 or 3 seems to be the inflection point, do the NMDS with 3 dimensions

#run the NMDS
# NMDS1 <- metaMDS(soil.D, k = 2, trymax = 100, trace = F)
# NMDS1
#not working because didn't include species names try again with species

#get the data frame
pcoa_env_df <- join(pcoa_df, sub.org.litter, type="left")

#do the nmds
NMDS2 <- metaMDS(pcoa_env_df[,c(14, 15, 16)], k = 2, trymax = 100, trace = F, autotransform = FALSE, distance = "bray")
NMDS2

plot(NMDS2)
text(NMDS2, display="species")

ordiplot(NMDS2)
orditorp(NMDS2, display="species")

#add the envfit
ef <- envfit(NMDS2, pcoa_env_df[ ,c(2, 3, 18, 19, 21, 22, 23)], permu=999)
ef

#plot the envfit
plot(NMDS2, type="t")
plot(ef, p.max = 0.05)
```


Does soil npoc vary by litter chemistry?
```{ r }
pcoa.npoc<-hhout[hhout$fractionated=="y", c("site", "species", "depth", "")]
pcoa.npoc
pcoa_env_df <- join(pcoa_env_df, pcoa.npoc, type = "left")

npocmod <- lm(npoc.mg.L ~ litter.pct.c, data = pcoa_env_df) 
npocmod2 <- lm(npoc.mg.L ~ litter.lignin.mg.g, data = pcoa_env_df)

summary(npocmod)
summary(npocmod2)
```


Okay, so npoc does vary with litter chemistry. Try adding the hwec.mg.g to the envfit
```{ r }

# do the nmds
# no convergent solutions with k=3, reduce to 2 dimensions
NMDS3 <- metaMDS(pcoa_env_df[,c(14, 15, 16, 22)], k = 2, trymax = 100, trace = F, autotransform = FALSE, distance = "bray")
NMDS3

plot(NMDS3)
text(NMDS3, display="species")

ordiplot(NMDS3)
orditorp(NMDS3, display="species")

#add the envfit - choose variables more selectively:
#  * depth
#  * lnratio
#  * species  - just for display purposes
  
ef <- envfit(NMDS3, pcoa_env_df[ ,c(2, 3, 19, 23)], permu=999)
ef

#plot the envfit
plot(NMDS3, type="t")
plot(ef, p.max = 0.05)
```







