---
title: "Plot_soil_data_spatially"
author: "Becca"
date: "22-5-2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r }
library(plotKML)
library(XML)
library(OpenStreetMap)
library(lubridate)
library(dplyr)
library(tidyverse)
library(vegan)
library(plyr)
library(ggplot2)
library(ggmap)
devtools::install_github("dkahle/ggmap", ref = "tidyup")

```


Load the combined fractionation data/location data df.
```{r }
cn.loc<-read.csv("cn.loc.csv", header=TRUE)
summary(cn.loc)
cn.loc$depth = as.factor(cn.loc$depth)

#rename all picea to 'picea' 
cn.loc$species<-mapvalues(cn.loc$species, from=c("picea_sit", "picea_ab"), to=c("picea", "picea"))
summary(cn.loc)
```


Get a stamen map:
```{r }
#set map boundaries
mylocation <- c(5.415, 52.450, 5.475, 52.483) 

#get the map from the stamen map database
hhout<-get_map(location=mylocation, source="stamen", maptype = "toner") 
str(hhout)

#pull out the lat/lon coordinates
#library(grDevices)
#coord<-[c(2,3)]
#coord<-as.matrix(coord)
```


Make the convex hulls dataframe for mapping later.
```{r }

#make the convex hulls based on the site coordinates
chulls<-ddply(cn.loc, .(species), 
              function(cn.loc) cn.loc[chull(cn.loc$lon, cn.loc$lat), ])
summary(chulls)
```


Make the basic plot:
```{r }
hhmap <- ggmap(hhout)
hhmap + 
  geom_point(data=cn.loc, mapping=aes(x=lon, y=lat, color=species)) +
  geom_polygon(data=chulls, aes(x=lon, y=lat, fill=species), alpha=0.3) +
      scale_color_brewer(palette="Set1") +
      scale_fill_brewer(palette="Set1") 
  
```


Try coloring points by bulk carbon and bulk nitrogen
```{r }
hhmap +
  geom_point(data=cn.loc[cn.loc$depth=="1",], mapping=aes(x=lon, y=lat, size=mean.c.pre, color=mean.c.post))  +
      scale_color_gradient2(low = "blue",
      high = "red", midpoint = 0.65)

hhmap +
  geom_point(data=cn.loc[cn.loc$depth=="1",], mapping=aes(x=lon, y=lat, size=mean.n.pre, color=mean.n.post)) + 
      scale_color_gradient2(low = "blue",
      high = "red", midpoint = 0.0036)
```


Color points by starting C, size by species and bulk pre-ashing c
```{r}
hhmap +
  geom_point(data=cn.loc, mapping=aes(x=lon, y=lat, size=mean.c.pre, color=species)) +
      scale_color_brewer(palette="Set1") +
      scale_fill_brewer(palette="Set1")
```


Try kriging to interpolate across variables.
```{r }
library(sp)
library(gstat)
library(scales)
library(magrittr)
library(sf)
library(raster)
library(automap)
```


Convert cn.loc into a spatial dataframe
```{r }
str(cn.loc)
c<-cn.loc
coordinates(c) <- ~lon + lat
class(c)
str(c)

#look at the bounding box
bbox(c)

#make grid
n <- c(50, 50)

cgrid<-
 st_make_grid(c, cellsize = c(diff(st_bbox(c)[c(1, 3)]),
  diff(st_bbox(c)[c(2, 4)]))/n, 
  offset = st_bbox(c)[c("xmin", "ymin")],
  crs = if (missing(x)) NA_crs_ else st_crs(c),
  what = "polygons", square = TRUE)
plot(cgrid)
cgrid <- as_Spatial(cgrid)
class(cgrid)

#kriging
kriging_result <- autoKrige(org.c.in.total.bulk ~ 1, cgrid)
plot(kriging_result)

```

          min       max
lat 52.454868 52.477868
lon  5.418579  5.452375



```{r }
if (!require("rspatial")) devtools::install_github('rspatial/rspatial')
library(rspatial)
c<-cn.loc
coordinates(c) <- ~lon + lat
class(c)

plot(sort(c$c.in.org.fraction.bulk))

csp<-SpatialPoints(cn.loc[, 32:33], proj4string=CRS("+proj=longlat +datum=NAD83"))
class(csp)
csp<-SpatialPointsDataFrame(csp)

csp <- SpatialPointsDataFrame(csp, c)

#make grid
n <- c(50, 50)

cgrid<-
 st_make_grid(c, cellsize = c(diff(st_bbox(c)[c(1, 3)]),
  diff(st_bbox(c)[c(2, 4)]))/n, 
  offset = st_bbox(c)[c("xmin", "ymin")],
  crs = if (missing(x)) NA_crs_ else st_crs(c),
  what = "centers", square = TRUE)
plot(cgrid)
class(cgrid)

cgrid2<-st_make_grid(c, cellsize = 0.1, what="centers")%>%
  st_intersection(c)

#kriging
kriging_result <- autoKrige(org.c.in.total.bulk ~ 1, c)
plot(kriging_result)

#make colors and polygons
cuts <- c(0, 0.15, 0.3, 0.45, 0.6, 0.75, 0.9, 1.05)
colors <-colorRampPalette(c('pink', 'yellow', 'orange', 'blue', 'dark blue'))
pols <- list(cgrid, fill = "lightgray")

spplot(csp, 'org.c.in.total.bulk', cuts=cuts, col.regions=colors(6), sp.layout=pols)
``` 

