---
title: "CN_wts_sums2_dataframe"
author: "Becca"
date: "10-7-2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Make a dataframe retaining the structure of the relationships we expect among variables.

Load libraries
```{r}
library(tidyverse)
library(plyr)
library(dplyr)
library(ggplot2)
```

Start by looking at the real data to generate hypothesized relationships
```{ r }
#first, load the cn data that we already have to generate data from 
cn.real.data<-read.csv("organic.c.values.csv", header=TRUE)
summary(cn.real.data)
cn.real.data$depth = as.factor(cn.real.data$depth)


carbon.plot.sp<-read.csv("carbon.plot.sp.csv", header=TRUE)
summary(carbon.plot.sp)
carbon.plot.sp$depth = as.factor(carbon.plot.sp$depth)
#get rid of the first two columns in the carbon.plot.sp dataframe
carbon.plot.sp<-carbon.plot.sp[ ,-c(1,2)]
summary(carbon.plot.sp)

#join the dataframes
cn.real.dat2<-join(cn.real.data, carbon.plot.sp, type="left")
summary(cn.real.dat2)
#totwt didn't work, delete it
cn.real.dat2<-cn.real.dat2[ ,-c(1,13)]
summary(cn.real.dat2)

#remake the carbon totwt variable
#lots of totwts missing. Redo this value
c.totwt<-aggregate(orgc.weight ~ site + depth, cn.real.dat2, sum) #total mass of organic C by site and depth
c.totwt$totwt<-c.totwt$orgc.weight
c.totwt<-c.totwt[,-(3)]
cn.real.dat3<-join(cn.real.dat2, c.totwt, type="left")
summary(cn.real.dat3)
#great. 
#Now make a sum-of-fractions variable
c.sums<-aggregate(weight.g ~ site + depth, cn.real.dat3, sum)
c.sums$final.wt<-c.sums$weight.g
c.sums<-c.sums[,-(3)]
cn.real.dat4<-join(cn.real.dat3, c.sums, type="left")
summary(cn.real.dat4)
#worked.

#get information on parameters of interest
summary(cn.real.dat2$soil.start.wt) #mean 10.47, min 10.02, max 11.45
hist(cn.real.dat2$org.c.in.t.frac, freq=FALSE) #bimodal, big weight btwn 0-5%, another peak ~25-30%
summary(cn.real.dat4$final.wt) #mean 10.550, min 7.876, max=11.791
summary(cn.real.dat4$totwt) #mean 0.476, min 0.1803, max 0.931
```


Get the data for the organic and inorganic C and N. 
```{ r }
cn.wts.sums<-read.csv("cn.wts.sums.csv", header=TRUE)
summary(cn.wts.sums)
cn.wts.sums$depth = as.factor(cn.wts.sums$depth)
#calculate starting c mass
cn.wts.sums$org.c.start.mass<-(cn.wts.sums$org.c.in.total.bulk/100)*cn.wts.sums$soil.start.wt #is this a valid way to measure this? Maybe we should correct for the change in soil weight from start to finish. 

#calculate c mass by fraction 
cn.wts.sums$org.c.fraction.mass<-(cn.wts.sums$weight.g*(cn.wts.sums$org.c.in.t.frac/100))

#calculate final (sum of fractions) c mass
cn.wts.sub <- cn.wts.sums %>%
  dplyr::group_by(site, depth, species) %>%
  dplyr::summarize(soil.final.wt = sum(weight.g),
            org.c.fraction.mass.sum = sum(org.c.fraction.mass), 
            )
summary(cn.wts.sub)

#try joining back with the main dataframe
cn.wts.sums2 <- join(cn.wts.sums, cn.wts.sub, type="left")
summary(cn.wts.sums2)
all.equal(cn.wts.sums2$totwt, cn.wts.sums2$org.c.fraction.mass.sum)

cn.wts.sums2$species<-mapvalues(cn.wts.sums2$species, from=c("picea_sit", "picea_ab"), to=c("picea", "picea"))
```


check out parameter values.
```{r}
summary(cn.wts.sums2$org.c.in.t.frac) #organic c pct. by fraction
summary(cn.wts.sums2$totwt) #eqal to org.c.fraction.mass.sum. 
hist(cn.wts.sums2$totwt) #Bimodal: peak between 0.2 and 0.4, another peak between 0.6 and 0.7
summary(cn.wts.sums2$orgc.weight) #weight of organic c per fraction
```


Translate % inorganic C to an inorganic C mass per fraction, and % inorganic C in bulk sample to inorganic C mass. Then plot the data.
```{r }
#inorganic fractions
cn.wts.sums2$inorg.c.mass.frac<-(cn.wts.sums2$pct.c.post.ashing/100)*(cn.wts.sums2$weight.g-((cn.wts.sums2$pct.mass.loss/100)*cn.wts.sums2$weight.g))


(inorg.c.frac<-ggplot(cn.wts.sums2, aes(x=species, y=inorg.c.mass.frac)) +
    geom_point()+
    geom_boxplot(alpha=0.3) +
    facet_grid(vars(depth), vars(fraction.code))+
    ylab("Inorganic C mass (g) by fraction and depth"))

#inorganic bulk
cn.wts.sums2$inorg.c.mass.bulk<-(cn.wts.sums2$mean.c.post/100)*cn.wts.sums2$soil.start.wt

(inorg.c.bulk<-ggplot(cn.wts.sums2, aes(x=species, y=inorg.c.mass.bulk)) +
    geom_point()+
    geom_boxplot(alpha=0.3) +
    facet_wrap(~depth) +
    ylab("Inorganic C mass (g) by species and depth"))
```


Plot the organic C data. 
```{r}

(hist.orgc<-hist(cn.wts.sums2$orgc.weight))

(orgc.plot.fraction<-ggplot(cn.wts.sums2, aes(x=species, y=orgc.weight)) +
                   geom_point(color=cn.wts.sums2$depth) +
                   geom_boxplot(alpha=0.3) +
                   facet_grid(vars(depth), vars(fraction.code))+
                   ylab("organic C mass (g) by fraction and depth"))

cn.wts.sums2$org.c.bulk.wt<-(cn.wts.sums2$org.c.in.total.bulk/100)*cn.wts.sums2$soil.start.wt

(hist.orgc.bulk<-hist(cn.wts.sums2$org.c.bulk.wt))

(org.c.plot.bulk<-ggplot(cn.wts.sums2, aes(x=species, y=org.c.bulk.wt)) +
                      geom_point(color=cn.wts.sums2$depth) +
                      geom_boxplot(alpha=0.3) +
                      facet_wrap(~depth) +
                      ylab("Organic C mass (g) in bulk sample by depth"))
```


Save the cn.wts.sums2 dataframe
```{r}
write.csv(cn.wts.sums2, file = "cn.wts.sums2.csv")
```
