---
title: "Master_dataframe"
author: "Becca"
date: "3/6/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

This is a file for harmonizing and combining data frames containing all of the data collected and processed for analyzing patterns in the Hollandse Hout field site. 

For the final data

The following variables should be included in the final data frame:

* litter chemistry
* soil organic/inorganic C
* soil fractions (pom, sand/agg fraction, silt/clay fraction, doc)
* soil pH
* soil moisture
* soil bulk density

UPDATE 30 APRIL: The most updated version of the hhout master dataframe has been saved as 'hhout_master.csv'. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r }
library(tidyverse)
#library(dplyr)
```

cn.wts.sums2
```{ r }
cn.wts.sums2 <- read.csv("cn.wts.sums2.csv")

summary(cn.wts.sums2)

cn.wts.sums2$depth <- as.factor(cn.wts.sums2$depth)
```


combined.data.cn
```{ r }
combined.data.cn <- read.csv("combined.data.cn.csv", header=TRUE)

summary(combined.data.cn)

combined.data.cn$depth <- as.factor(combined.data.cn$depth)

#rename the levels of species to match the litter_chem data
combined.data.cn$species <- recode(combined.data.cn$species, acer='acer', fagus='fagus', fraxinus='fraxinus', picea='picea', picea_ab='picea', picea_sit='picea', populus='populus', quercus='quercus')

levels(combined.data.cn$species)

summary(combined.data.cn)

#rename mis-labeled levels
combined.data.cn$site <- recode(combined.data.cn$site, '127G'= '157G', '130E' = '130E' ,'130E1' = '130E')

levels(combined.data.cn$site)
```


Remove levels of combined.data.cn that don't match with other data frames (there has to be a better way to do this).
```{ r }
#remove levels that don't match other data frames
#c("133V2", "157M", "130N1", "blank")

# newiris <- subset(iris, iris$Species != "virginica", drop = TRUE)
combined.data.cn.drop <- subset(combined.data.cn, combined.data.cn$site != "133V2", drop=TRUE)
# newiris$Species <- droplevels(newiris$Species) ## equivalent to newiris$Species[,drop=TRUE]
combined.data.cn.drop$site <- droplevels(combined.data.cn.drop$site)

combined.data.cn.drop <- subset(combined.data.cn.drop, combined.data.cn.drop$site != "157M", drop=TRUE)
combined.data.cn.drop$site <- droplevels(combined.data.cn.drop$site)

combined.data.cn.drop <- subset(combined.data.cn.drop, combined.data.cn.drop$site != "130N1", drop=TRUE)
combined.data.cn.drop$site <- droplevels(combined.data.cn.drop$site)

combined.data.cn.drop <- subset(combined.data.cn.drop, combined.data.cn.drop$site != "blank", drop=TRUE)
combined.data.cn.drop$site <- droplevels(combined.data.cn.drop$site)

levels(combined.data.cn.drop$site)


```

Rename this awful-long dataframe
```{ r }
cn.dat <- combined.data.cn.drop
write.csv(cn.dat, file="cn.dat.csv", row.names=FALSE)
```


Read in the new data frame
```{ r }
cn.dat<-read.csv("cn.dat.csv", header=TRUE)

summary(cn.dat)
cn.dat$depth <- as.factor(cn.dat$depth)
summary(cn.dat)

levels(cn.dat$site)

```

Combine the sites that have been arbitrarily separated
```{ r }
#site 130 depth 1
cn.dat$mean.c.pre[c(27)] <- cn.dat$mean.c.pre[c(138)]
cn.dat$mean.n.pre[c(27)] <- cn.dat$mean.n.pre[c(138)]
cn.dat$mean.c.post[c(27)] <- cn.dat$mean.c.post[c(138)]
cn.dat$mean.n.post[c(27)] <- cn.dat$mean.n.post[c(138)]
cn.dat$mean.pct.mass.loss[c(27)] <- cn.dat$mean.pct.mass.loss[c(138)]
cn.dat[c(27),]

#site 130 depth 2
cn.dat$mean.c.pre[c(28)] <- cn.dat$mean.c.pre[c(139)]
cn.dat$mean.n.pre[c(28)] <- cn.dat$mean.n.pre[c(139)]
cn.dat$mean.c.post[c(28)] <- cn.dat$mean.c.post[c(139)]
cn.dat$mean.n.post[c(28)] <- cn.dat$mean.n.post[c(139)]
cn.dat$mean.pct.mass.loss[c(28)] <- cn.dat$mean.pct.mass.loss[c(139)]
cn.dat[c(28),]

#site 130 depth 3
cn.dat$mean.c.pre[c(29)] <- cn.dat$mean.c.pre[c(140)]
cn.dat$mean.n.pre[c(29)] <- cn.dat$mean.n.pre[c(140)]
cn.dat$mean.c.post[c(29)] <- cn.dat$mean.c.post[c(140)]
cn.dat$mean.n.post[c(29)] <- cn.dat$mean.n.post[c(140)]
cn.dat$mean.pct.mass.loss[c(29)] <- cn.dat$mean.pct.mass.loss[c(140)]
cn.dat[c(29),]

#site 130 depth 4
cn.dat$mean.c.pre[c(30)] <- cn.dat$mean.c.pre[c(141)]
cn.dat$mean.n.pre[c(30)] <- cn.dat$mean.n.pre[c(141)]
cn.dat$mean.c.post[c(30)] <- cn.dat$mean.c.post[c(141)]
cn.dat$mean.n.post[c(30)] <- cn.dat$mean.n.post[c(141)]
cn.dat$mean.pct.mass.loss[c(30)] <- cn.dat$mean.pct.mass.loss[c(141)]
cn.dat[c(30),]

View(cn.dat)
```

Drop unnecessary rows
```{ r }
cn.dat<-cn.dat[-c(138, 139, 140, 141), ]

View(cn.dat)
```

cn.loc
```{ r }
cn.loc <- read.csv("cn.loc.csv")
summary(cn.loc)

#make depth a factor
cn.loc$depth <- as.factor(cn.loc$depth)

#rename the levels of species to match the litter_chem data
cn.loc$species <- recode(cn.loc$species, acer='acer', fagus='fagus', fraxinus='fraxinus', picea_ab='picea', picea_sit='picea', populus='populus', quercus='quercus')

levels(cn.loc$species)
```


litter chemistry
```{ r }
litter_chem <- read.csv("litter_chem.csv")

summary(litter_chem)

litter_chem <- 
litter_chem %>%
  rename(
    litter.pct.n = pct.n, 
    litter.pct.c = pct.c, 
    litter.lignin.mg.g = lignin.mg.g,
    litter.cell.mg.g = cell.mg.g, 
    litter.hwec.ppm = mean_hwec_ppm
  )
```


Check concordance among dataframes using common variables
```{ r }
all.equal(cn.wts.sums2$site, cn.loc$site)
#true
all.equal(cn.loc$site, cn.dat$site)
#very different. 

levels(cn.dat$site)
levels(cn.loc$site)
#this is because combined.data.cn has more levels than the other data frames.
#extra sites in cn.dat are: 131V2, 156Q, and 160D

all.equal(cn.wts.sums2$pct.n.pre.ashing, cn.loc$pct.n.pre.ashing)
#okay, so these are the same
```


Check what variables are in each data frame
```{ r }
names(cn.loc)
```
> names(cn.loc)
 [1] "X"                      "site"                   "depth"                  "fraction.name"         
 [5] "fraction.code"          "pct.n.pre.ashing"       "pct.c.pre.ashing"       "pct.n.post.ashing"     
 [9] "pct.c.post.ashing"      "pct.mass.loss"          "c.in.org.fract"         "org.c.in.t.frac"       
[13] "n.in.org.fract"         "org.n.in.t.frac"        "species"                "mean.c.pre"            
[17] "mean.n.pre"             "mean.c.post"            "mean.n.post"            "mean.pct.mass.loss"    
[21] "fraction.id"            "c.in.org.fraction.bulk" "org.c.in.total.bulk"    "n.in.org.fraction.bulk"
[25] "org.n.in.total.bulk"    "sample"                 "soil.start.wt"          "weight.g"              
[29] "orgc.weight"            "totwt"                  "pct.c.cont.by.fraction" "lon"                   
[33] "lat"                    "ele"                    "time"                   "name"

#the variables we need are everything except for X, ele, time, and name

```{ r}
names(cn.dat)
```

#we need everything except level 

```{ r}
names(cn.wts.sums2)
```

> names(cn.wts.sums2)
 [1] "site"                    "depth"                   "fraction.name"           "fraction.code"          
 [5] "pct.n.pre.ashing"        "pct.c.pre.ashing"        "pct.n.post.ashing"       "pct.c.post.ashing"      
 [9] "pct.mass.loss"           "c.in.org.fract"          "org.c.in.t.frac"         "n.in.org.fract"         
[13] "org.n.in.t.frac"         "species"                 "mean.c.pre"              "mean.n.pre"             
[17] "mean.c.post"             "mean.n.post"             "mean.pct.mass.loss"      "fraction.id"            
[21] "c.in.org.fraction.bulk"  "org.c.in.total.bulk"     "n.in.org.fraction.bulk"  "org.n.in.total.bulk"    
[25] "sample"                  "soil.start.wt"           "weight.g"                "orgc.weight"            
[29] "totwt"                   "pct.c.cont.by.fraction"  "org.c.start.mass"        "org.c.fraction.mass"    
[33] "soil.final.wt"           "org.c.fraction.mass.sum" "inorg.c.mass.frac"       "inorg.c.mass.bulk"      
[37] "org.c.bulk.wt"     

#use everything 

```{ r}
names(litter_chem)
```

#use everything

Join cn.wts.sums2 and cn.loc
```{ r }
#first subset the cn.loc values that I don't need (X, ele, time, and name)
cn.loc.sub <- cn.loc[c(-1,-34, -35, -36)]
names(cn.loc.sub)

#now do a full join
join1 <- full_join(cn.wts.sums2, cn.loc.sub)

#check summary
summary(join1)
```

Joining, by = c("site", "depth", "fraction.name", "fraction.code", "pct.n.pre.ashing", "pct.c.pre.ashing", "pct.n.post.ashing", "pct.c.post.ashing", "pct.mass.loss", "c.in.org.fract", "org.c.in.t.frac", "n.in.org.fract", "org.n.in.t.frac", "species", "mean.c.pre", "mean.n.pre", "mean.c.post", "mean.n.post", "mean.pct.mass.loss", "fraction.id", "c.in.org.fraction.bulk", "org.c.in.total.bulk", "n.in.org.fraction.bulk", "org.n.in.total.bulk", "sample", "soil.start.wt", "weight.g", "orgc.weight", "totwt", "pct.c.cont.by.fraction")

Now join the cn.dat
```{ r }
#remove the values of cn.dat I don't need (depth_cm)
cn.dat.sub <- cn.dat[c(-4)]
names(cn.dat.sub)

#now do a full join
join2 <- full_join(join1, cn.dat.sub)
```

Joining, by = c("site", "depth", "species", "mean.c.pre", "mean.n.pre", "mean.c.post", "mean.n.post", "mean.pct.mass.loss")

* everything in the cn.dat.sub that had a match with the join1 df worked!

Make site and depth factors again
```{ r }
join2$site <- as.factor(join2$site)
join2$depth <- as.factor(join2$depth)
```

Now join with the litter data
```{ r}
join3 <- full_join(join2, litter_chem)

summary(join3)
head(join3)
```

Yay! Complete data frame!

Now save as a Master dataframe so we can use it in analysis
```{ r }
join3$depth <- as.factor(join3$depth)

write.csv(join3, file = "MasterDF.csv", row.names = FALSE)
```

Read in the dataframe to check it works
```{ r }
masterdf <- read.csv("MasterDF.csv", header=TRUE)

summary(masterdf)
masterdf$depth <- as.factor(masterdf$depth)
masterdf$site <- as.factor(masterdf$site)
```

### 06 March 2020 

Okay, so some sites have discrepancies in the data. Sites to check: 

a. 126E - discrepancy in mean.c.pre, mean.n.pre, mean.c.post, and mean.n.post for *depth 1*
b. 126F - *very large* discrepancy in the mean.c.pre, mean.n.pre, mean.c.post, mean.n.post, and mean.pct.mass.loss in *depth 2*
c. 127F - *very small* discrepancy in mean.c.pre, mean.n.pre, mean.c.post, mean.n.post, and mean.pct.mass.loss in *depth 2*
d. 131N1 - *moderate* discrepancy in mean.c.pre, mean.n.pre, mean.c.post, mean.n.post, and mean.pct.mass.loss in *depth 1* and *depth 2*
e. 131N2 - *mistake in decimal placement* for *mean.c.post* (decimal moved 1 place to the left) for *depth 2*
f. 131V1 - *moderate discrepancy* in mean.c.pre, mean.n.pre, mean.c.post, mean.n.post, and mean.pct.mass.loss in *depth 2*
g. 131V2 - this was a site measured in round 1 that was *not measured in fractionation*
h. 132Q - *large* discrepancy in  mean.c.pre, mean.n.pre, mean.c.post, mean.n.post, and mean.pct.mass.loss in *depth 2* - I think what I have for depth 2 in the soil fractions section *might actually be depth 3*!!
i. 134T - *small* discrepancy in mean.c.pre, mean.n.pre, mean.c.post, mean.n.post, and mean.pct.mass.loss in *depth 1* and *depth 2* 
j. 134X - *moderate* discrepancy in mean.c.pre, mean.n.pre, mean.c.post, mean.n.post, and mean.pct.mass.loss in *depth 2*
k. 135J - *small* discrepancy in mean.c.pre, mean.n.pre, mean.c.post, mean.n.post, and mean.pct.mass.loss in *depth 2*
l. 156Q - this was another site measured in round 1 that was *not measured in fractionation*
m. 157G - *no apparent discrepancy*, not sure why they aren't matching up -possibly because of the *NAs that are being written as <NA> ?*
n. 159L - *small* discrepancy in mean.c.pre, mean.n.pre, mean.c.post, mean.n.post, and mean.pct.mass.loss in *depth 1* and *depth 2*
o. 160D - this was another site measured in round 1 that was not measured in fractionation

There are discrepancies because I averaged the values of the mean.c.pre, mean.n.pre, mean.c.post, and mean.n.post for the bulk soil dataframe but did not do so for the cn.wts.sums dataframe. 

### 09 March 2020

## Correct simple errors first

* 157G - added the pH from the bulk df to the fractionation df and deleted the rest of the row
* 131N2 - moved the decimal place for mean.c.post

adding mean.ph, avg.bd, pct.m from row 216 to rows 73, 74, 75 (131N2 depth 2)
```{ r}
#deck2$value[c(13, 26, 39, 52)] <- 14
masterdf$mean.ph[c(7, 8, 9)] <- masterdf$mean.ph[c(188)]
masterdf$avg.bd[c(7, 8, 9)] <- masterdf$avg.bd[c(188)]
masterdf$pct.m[c(7, 8, 9)] <- masterdf$pct.m[c(188)]
masterdf[c(7, 8, 9),]

#check the 126E
masterdf[masterdf$site=="126E", ]
```


Check 157G
```{ r }
masterdf[masterdf$site=='157G', ]

#add the pH from the bulk sites to fracitonated samples
masterdf$mean.ph[c(130, 131, 132)] <- masterdf$mean.ph[c(203)]
masterdf$mean.ph[c(133, 134, 135)] <- masterdf$mean.ph[c(204)]

masterdf[masterdf$site=='157G', ]
```

Can't seem to get rid of that last na row, leave it for now.

Check 126F
```{ r }
masterdf[masterdf$site=='126F', ]

#take the average of mean.c.pre, mean.c.post, mean.n.pre, mean.n.post for depth 2 (193 to 13, 14, 15)
masterdf$mean.c.pre[c(13, 14, 15)] <- mean(masterdf$mean.c.pre[c(193, 13)])
masterdf$mean.n.pre[c(13, 14, 15)] <- mean(masterdf$mean.n.pre[c(193, 13)])
masterdf$mean.c.post[c(13, 14, 15)] <- mean(masterdf$mean.c.post[c(193, 13)])
masterdf$mean.n.post[c(13, 14, 15)] <- mean(masterdf$mean.n.post[c(193, 13)])
masterdf$mean.pct.mass.loss[c(13, 14, 15)] <- mean(masterdf$mean.pct.mass.loss[c(13, 193)])
masterdf[c(13, 14, 15),]

#move mean.ph to depths 1 and 2 fractionation (192 to 10, 11, 12 and 193 to 13, 14, 15)
masterdf$mean.ph[c(13, 14, 15)] <- masterdf$mean.ph[c(193)]
masterdf$mean.ph[c(10, 11, 12)] <- masterdf$mean.ph[c(192)]

masterdf[masterdf$site=='126F', ]
#done
```

Check 127F
```{ r }
masterdf[masterdf$site=='127F', ]

#take the average of mean.c.pre, mean.c.post, mean.n.pre, mean.n.post for depth 2 (193 to 13, 14, 15)
masterdf$mean.c.pre[c(31, 32, 33)] <- mean(masterdf$mean.c.pre[c(200, 31)])
masterdf$mean.n.pre[c(31, 32, 33)] <- mean(masterdf$mean.n.pre[c(200, 31)])
masterdf$mean.c.post[c(31, 32, 33)] <- mean(masterdf$mean.c.post[c(200, 31)])
masterdf$mean.n.post[c(31, 32, 33)] <- mean(masterdf$mean.n.post[c(200, 31)])
masterdf$mean.pct.mass.loss[c(31, 32, 33)] <- mean(masterdf$mean.pct.mass.loss[c(200, 31)])
masterdf[c(31, 32, 33),]

#move mean.ph to depths 1 and 2 fractionation (192 to 10, 11, 12 and 193 to 13, 14, 15)
masterdf$mean.ph[c(28, 29, 30)] <- masterdf$mean.ph[c(199)]
masterdf$mean.ph[c(31, 32, 33)] <- masterdf$mean.ph[c(200)]

masterdf[masterdf$site=='127F', ]
#done
```

Check 131N1
```{ r }
masterdf[masterdf$site=='131N1', ]

#move mean.ph & avg.bd to depths 1 and 2 fractionation (217 to 64:66) and 
masterdf$mean.ph[c(64, 65, 66)] <- masterdf$mean.ph[c(217)]
masterdf$avg.bd[c(64, 65, 66)] <- masterdf$mean.ph[c(217)]
#218 to 67:69
masterdf$mean.ph[c(67, 68, 69)] <- masterdf$mean.ph[c(218)]
masterdf$avg.bd[c(67, 68, 69)] <- masterdf$avg.bd[c(218)]


masterdf[masterdf$site=='131N1', ]
#done
```


Check 131N2
```{ r }
masterdf[masterdf$site=='131N2', ]
#move mean.pg, avg.bd, and pct.m from 221 to 73, 74, 75
masterdf$mean.ph[c(73, 74, 75)] <- masterdf$mean.ph[c(221)]#7.3580
masterdf$avg.bd[c(73, 74, 75)] <- masterdf$avg.bd[c(221)]#0.978
masterdf$pct.m[c(73, 74, 75)] <- masterdf$pct.m[c(221)]#24.489

masterdf[masterdf$site=='131N2', ]
```


Check 131V1
```{ r }
masterdf[masterdf$site=='131V1', ]

#move depth 2 ph, bd and pct.m to 79-81
masterdf$mean.ph[c(79, 80, 81)] <- masterdf$mean.ph[c(224)]#7.322
masterdf$avg.bd[c(79, 80, 81)] <- masterdf$avg.bd[c(224)]#0.358
masterdf$pct.m[c(79, 80, 81)] <- masterdf$pct.m[c(224)]#18.439

masterdf[masterdf$site=='131V1', ]
#done
```


Check 132Q
```{ r }
masterdf[masterdf$site=='132Q', ]
#depth 1
masterdf$mean.ph[c(82, 83, 84)] <- masterdf$mean.ph[c(85)]
masterdf$avg.bd[c(82, 83, 84)] <- masterdf$avg.bd[c(85)]
masterdf$pct.m[c(82, 83, 84)] <- masterdf$pct.m[c(85)]
#depth 2
masterdf$mean.ph[c(85, 86, 87)] <- masterdf$mean.ph[c(231)]
masterdf$avg.bd[c(85, 86, 87)] <- masterdf$avg.bd[c(231)]
masterdf$pct.m[c(85, 86, 87)] <- masterdf$pct.m[c(231)]

masterdf[masterdf$site=='132Q', ]
#ok.
```


Check 134T
```{ r }
masterdf[masterdf$site=='134T', ]

#move depth 1 ph, bd and pct.m to 106-108
masterdf$mean.ph[c(106, 107, 108)] <- masterdf$mean.ph[c(240)]#6.908
masterdf$avg.bd[c(106, 107, 108)] <- masterdf$avg.bd[c(240)]#0.739
masterdf$pct.m[c(106, 107, 108)] <- masterdf$pct.m[c(240)]#66.555

#move depth 2 ph, bd and pct.m to 109-111
masterdf$mean.ph[c(109, 110, 111)] <- masterdf$mean.ph[c(241)]#7.224
masterdf$avg.bd[c(109, 110, 111)] <- masterdf$avg.bd[c(241)]#0.904
masterdf$pct.m[c(109, 110, 111)] <- masterdf$pct.m[c(241)]#56.755

masterdf[masterdf$site=='134T', ]

#ok
```


Check 134X
```{ r }
masterdf[masterdf$site=='134X', ]

#take the average of mean.c.pre, mean.c.post, mean.n.pre, mean.n.post for depth 2 (193 to 13, 14, 15)
# masterdf$mean.c.pre[c(115, 116, 117)] <- mean(masterdf$mean.c.pre[c(244, 115)])
# masterdf$mean.n.pre[c(115, 116, 117)] <- mean(masterdf$mean.n.pre[c(244, 115)])
# masterdf$mean.c.post[c(115, 116, 117)] <- mean(masterdf$mean.c.post[c(244, 115)])
# masterdf$mean.n.post[c(115, 116, 117)] <- mean(masterdf$mean.n.post[c(244, 115)])
# masterdf$mean.pct.mass.loss[c(115, 116, 117)] <- mean(masterdf$mean.pct.mass.loss[c(244, 115)])
# masterdf[c(115, 116, 117),]

#move mean.ph to depths 1 and 2 fractionation 
masterdf$mean.ph[c(115, 116, 117)] <- masterdf$mean.ph[c(244)]
masterdf$avg.bd[c(115, 116, 117)] <- masterdf$avg.bd[c(244)]
masterdf$pct.m[c(115, 116, 117)] <- masterdf$pct.m[c(244)]

masterdf[masterdf$site=='134X', ]
```

Check 135J
```{ r }
masterdf[masterdf$site=='135J', ]

masterdf$mean.c.pre[c(118, 119, 120)] <- mean(masterdf$mean.c.pre[c(118, 247)])
masterdf$mean.n.pre[c(118, 119, 120)] <- mean(masterdf$mean.n.pre[c(118, 247)])
masterdf$mean.c.post[c(118, 119, 120)] <- mean(masterdf$mean.c.post[c(118, 247)])
masterdf$mean.n.post[c(118, 119, 120)] <- mean(masterdf$mean.n.post[c(118, 247)])
masterdf$mean.pct.mass.loss[c(118, 119, 120)] <- mean(masterdf$mean.pct.mass.loss[c(118, 247)])

masterdf$mean.ph[c(118, 119, 120)] <- masterdf$mean.ph[c(247)]
masterdf$avg.bd[c(118, 119, 120)] <- masterdf$avg.bd[c(247)]
masterdf$pct.m[c(118, 119, 120)] <- masterdf$pct.m[c(247)]

masterdf[masterdf$site=='135J', ]
```

Check 156Q
```{ r }
masterdf[masterdf$site=='156Q', ]
#not present in fractionation sample
```

Check 159L
```{ r }
masterdf[masterdf$site=='159L', ]

masterdf$mean.ph[c(164, 165, 166)] <- masterdf$mean.ph[c(265)]
masterdf$avg.bd[c(164, 165, 166)] <- masterdf$avg.bd[c(265)]
masterdf$pct.m[c(164, 165, 166)] <- masterdf$pct.m[c(265)]

masterdf[masterdf$site=='159L', ]
```

Check160D
```{ r }
masterdf[masterdf$site=='160D', ]
#not present in fractionation sample
```


Save the masterdf dataframe
```{ r }
write.csv(masterdf, file = "masterdf.csv", row.names = FALSE)
```


Remove the extra rows
```{ r }
#find the rows that are just the bulk soil sample
View(masterdf)

#Drop extra rows
masterdf.complete.cases <- masterdf[-c(188, 189, 192, 193, 199, 200, 203, 204, 214, 217, 218, 221, 224, 231, 240, 241, 244, 247, 262, 263, 265, 266), ] 

View(masterdf.complete.cases)
#good
```


Check for missing values
```{ r }
summary(masterdf.complete.cases)

#4 rows with missing species, check those
masterdf.complete.cases[masterdf.complete.cases$site=="157G", ]
#row 205 = populus
masterdf.complete.cases[masterdf.complete.cases$site=="126F", ] 
#rows 194 & 280 = picea
masterdf.complete.cases[masterdf.complete.cases$site=="131N1", ] 
#row 220 = fagus

#replacing these by hand, I'm so frustrated by the R code right now. 
write.csv(masterdf.complete.cases, file = "masterdf.complete.csv", row.names=FALSE)
```

Replace missing species values by hand.

Read in the 'fixed' dataframe
```{ r }
master <- read.csv("masterdf.complete.csv", header = TRUE)
summary(master)
master$depth <- as.factor(master$depth)
```


Now all of the species are accounted for, but some of the litter values are missing because I joined these dataframes before fixing the species issue. Go ahead and fill these in with the correct values
```{ r }
master[master$site=='157G', ]

master$litter.pct.n[c(197)] <- master$litter.pct.n[c(135)]
master$litter.pct.c[c(197)] <- master$litter.pct.c[c(135)]
master$litter.lignin.mg.g[c(197)] <- master$litter.lignin.mg.g[c(135)]
master$litter.cell.mg.g[c(197)] <- master$litter.cell.mg.g[c(135)]
master$litter.hwec.ppm[c(197)] <- master$litter.hwec.ppm[c(135)]

master[master$site=='157G', ]
```


```{ r }
master[master$site=='126F', ]

master$litter.pct.n[c(190, 258)] <- master$litter.pct.n[c(15)]
master$litter.pct.c[c(190, 258)] <- master$litter.pct.c[c(15)]
master$litter.lignin.mg.g[c(190, 258)] <- master$litter.lignin.mg.g[c(15)]
master$litter.cell.mg.g[c(190, 258)] <- master$litter.cell.mg.g[c(15)]
master$litter.hwec.ppm[c(190, 258)] <- master$litter.hwec.ppm[c(15)]

master[master$site=='126F', ]
```


```{ r }
master[master$site=='131N1', ]

master$litter.pct.n[c(209)] <- master$litter.pct.n[c(208)]
master$litter.pct.c[c(209)] <- master$litter.pct.c[c(208)]
master$litter.lignin.mg.g[c(209)] <- master$litter.lignin.mg.g[c(208)]
master$litter.cell.mg.g[c(209)] <- master$litter.cell.mg.g[c(208)]
master$litter.hwec.ppm[c(209)] <- master$litter.hwec.ppm[c(208)]

master[master$site=='131N1', ]
```


Check df
```{r}
summary(master)
```

All of the litter NAs are gone, all of the species NAs are gone, the rest are either missing values (ie not recorded) or are from depths 3 and 4, where there is no fractionation data. 

Add the NO3 and NH4 data
```{ r }
no3nh4<-read.csv2("no4_no3_bulk_field_data.csv", header=TRUE)
summary(no3nh4)

#get rid of extra rows
no3nh4 <- no3nh4[ , 1:6]
summary(no3nh4)

no3nh4$depth <- as.factor(no3nh4$depth)
summary(no3nh4)

#get rid of the cup nr and depth_cm
no3nh4 <- no3nh4[ , c(1,3,5,6)]
summary(no3nh4)
```

Combine with the master df
```{ r }
#check levels of no3nh4$site and levels of master$site
levels(master$site)
levels(no3nh4$site)
#mismatches on:
#  127G (only in no3nh4 data);
#  130Q1 (called 130Qin no3nh4);
#  130T2 (called 130T in no3nh4);
#  131N1 & 131N2 (called 130T in no3nh4);
#  131V1 & 131V2 (missing in no3nh4);
#  132Q (called 131Q in no3nh4);
#  

master.with.no3 <- full_join()
```

## NOPE

Not going to add this because it's hopelessly mislabeled. There's no way to tell which label goes with which sample. Thanks, Rob.


Add the npoc data to this
```{ r }
npoc <- read.csv("npoc_hhout.csv", header=TRUE)
summary(npoc)
npoc$depth <- as.factor(npoc$depth)
```

Subset out the columns I want (depth, site, species, npoc)
```{ r }
npoc.sub <- npoc[ ,c(2,3,4,7)]
summary(npoc.sub)
#View(npoc.sub)

#take out the control samples
npoc.sub <- na.omit(npoc.sub)
summary(npoc.sub)

#subset out the control data
#newiris <- subset(iris, iris$Species != "virginica", drop = TRUE)
npoc.sub <- subset(npoc.sub, npoc.sub$species != "control", drop = TRUE)
npoc.sub$species <- droplevels(npoc.sub$species)

summary(npoc.sub)
levels(npoc.sub$species)
```

Combine the npoc data with the master data frame
```{ r }
library(dplyr)
#check levels
levels(npoc.sub$site)
levels(master$site)

#do a full join
join4 <- full_join(master, npoc.sub)
summary(join4)

#make site and depth factors again
join4$site <- as.factor(join4$site)
join4$depth <- as.factor(join4$depth)

#add the units to the npoc variable
join4 <- 
  join4 %>% rename(npoc.mg.L = npoc)

summary(join4)
```


Save the combined dataframe
```{ r}
write.csv(join4, file = "hhout_master.csv", row.names = FALSE)
```

Remove the extra/duplicated variables
```{ r}
library(dplyr)
#read in the dataframe
hhout <- read.csv("hhout_master.csv", header=TRUE)

#drop the fraction.id variable, not needed
hhout <- select(hhout, -"fraction.id")
names(hhout)

#drop 'sample' - not needed since it duplicates the information in 'site'
hhout <- select(hhout, -"sample")
names(hhout)

#check that totwt and org.c.fraction.mass.sum are the same, then drop org.c.bulk.wt
all.equal(hhout$totwt, hhout$org.c.fraction.mass.sum)
#true, drop totwt
hhout <- select(hhout, -"totwt")
names(hhout)

#rename org.c.bulk.wt to org.c.mass.bulk 
hhout <- rename(hhout, org.c.mass.bulk = org.c.bulk.wt)
names(hhout)

#drop n, not relevant for data analysis
hhout <- select(hhout, -"n")
names(hhout)

```


save the updated dataframe
```{ r}
hhout_master_drop <- write.csv(hhout, file = "hhout_master_drop.csv", row.names = FALSE)
```


All duplicated/unnecessary rows have been removed. Now rename the non-informative variable names to make them more transparent.
```{ r }
#read in the updated dataframe
hhout <- read.csv("hhout_master_drop.csv", header=TRUE)

#1-4 are ok
#5-9 add 'fraction' at the end
names(hhout)[names(hhout) == "pct.n.pre.ashing"] <- "pct.n.pre.ashing.fraction" 
  
names(hhout)[names(hhout) == "pct.n.post.ashing"] <- "pct.n.post.ashing.fraction" 

names(hhout)[names(hhout) == "pct.c.pre.ashing"] <- "pct.c.pre.ashing.fraction" 

names(hhout)[names(hhout) == "pct.c.post.ashing"] <- "pct.c.post.ashing.fraction" 

names(hhout)

names(hhout)[names(hhout)== "pct.mass.loss"] <- "pct.mass.loss.fraction"

names(hhout)

#10-13 add pct to indicate these are percentages
names(hhout)[names(hhout) == "c.in.org.fract"] <- "pct.c.in.org.fraction" 

names(hhout)[names(hhout) == "n.in.org.fract"] <- "pct.n.in.org.fraction" 

names(hhout)[names(hhout) == "org.c.in.t.frac"] <- "pct.c.in.whole.fraction" 

names(hhout)[names(hhout) == "org.n.in.t.frac"] <- "pct.n.in.whole.fraction" 

names(hhout)

#15-18 change to 'pct' and 'pre' vs 'post' ashing 'bulk'
names(hhout)[names(hhout) == "mean.c.pre"] <- "pct.c.pre.ashing.bulk" 

names(hhout)[names(hhout) == "mean.c.post"] <- "pct.c.post.ashing.bulk" 

names(hhout)[names(hhout) == "mean.n.pre"] <- "pct.n.pre.ashing.bulk" 

names(hhout)[names(hhout) == "mean.n.post"] <- "pct.n.post.ashing.bulk" 

names(hhout)

#19-23
names(hhout)[names(hhout) == "mean.pct.mass.loss"] <- "pct.mass.loss.bulk" 

names(hhout)[names(hhout) == "c.in.org.fraction.bulk"] <- "pct.c.in.org.fraction.bulk" 

names(hhout)[names(hhout) == "org.n.in.total.bulk"] <- "pct.n.in.total.bulk" 

names(hhout)[names(hhout) == "n.in.org.fraction.bulk"] <- "pct.n.in.org.fraction.bulk" 

names(hhout)

#24-27 change 'weight' to 'mass' for accuracy 
names(hhout)[names(hhout) == "soil.start.wt"] <- "soil.mass.pre.fract" 

names(hhout)[names(hhout)=="weight.g"] <- "fraction.mass"

names(hhout)[names(hhout)=="orgc.weight"] <- "org.c.mass.fraction"

names(hhout)[names(hhout)=="pct.c.cont.by.fraction"] <- "pct.c.in.fraction"

names(hhout)


#28-31 specify bulk or fraction, change 'wt' to 'mass', 37-38 rewrite, 45 add 'soil'
names(hhout)[names(hhout)=="org.c.start.mass"] <- "org.c.start.mass.bulk"

names(hhout)[names(hhout)=="orc.c.fraction.mass"] <- "org.c.mass.fraction"

names(hhout)[names(hhout)=="soil.final.wt"] <- "soil.mass.sum.fractions"

names(hhout)[names(hhout)=="org.c.fraction.mass.sum"] <- "sum.org.c.mass.fraction"

names(hhout)

names(hhout)[names(hhout)=="mean.ph"] <- "soil.ph"

names(hhout)[names(hhout)=="avg.bd"] <- "bulk.dens"

names(hhout)[names(hhout)=="npoc.mg.L"] <- "soil.npoc.mg.L"

names(hhout)
```


Save the updated data frame
```{ r}
write.csv(hhout, file = "hhout_master_17Mar2020", row.names = FALSE)
```


Read in and check the updated data frame
```{ r }
hhout <- read.csv("hhout_master_17Mar2020", header=TRUE)
names(hhout)
summary(hhout)

#make depth a factor
hhout$depth = as.factor(hhout$depth)
```

Everything looks good!

I will keep the CUE data separate since it only applies to the top layer of the soil data.
