---
title: "Bayes_soil_C"
author: "Becca"
date: "10-5-2019"
output: pdf_document
  toc: true
editor_options: 
  chunk_output_type: console
---

#Bayesian modeling to constrain inorganic variance. 

The goal of this process is to constrain the range of inorganic C loss to within a plausible range. At the moment, an underestimation of inorganic C post ashing is leading to over-estimation of soil C content.  

##Describing the model

Model : We expect the output, inorganic C loss, to be a gamma-distributed (or other right-skewed distribution) with a lower limit of 0 and upper limit of 1 (100% mass loss).
      : The 

**Load required libraries
```{r }
#libraries
library(rstanarm)
library(brms)
library(bayesplot)
library(ggplot2)
#library(plyr)
library(dplyr)
library(tidybayes)
library(modelr)

```


**Make the data set that will be read into the model**
```{r }
c_data <- read.csv("cn.wts.sums.csv")
summary(c_data)
c_data$depth <- as.factor(c_data$depth)

#select just the relevant parameters
c_model_data <- subset(c_data, select = c(site, depth, fraction.code, mean.c.pre, mean.c.post, pct.c.pre.ashing, pct.c.post.ashing, weight.g, soil.start.wt, mean.pct.mass.loss, pct.mass.loss))
summary(c_model_data)
#c_model_data$depth <- as.factor(c_model_data$depth)

#calculate the bulk total c mass pre- and post- ashing and the sum of fractions total c mass pre- and post-ashing
#####AGGREGATE#EXAMPLE###############
# soil.loss.df1<-soil.loss.df %>% 
#   group_by(site, depth, species) %>%
#   summarize(
#     soil.start.wt = mean(soil.start.wt),
#     weight.g = sum(weight.g)) 
# soil.loss.df1
####################################

#summarize into the parameters relevant for this model
cmod_dat<-c_model_data %>%
  group_by(site, depth) %>%
  summarize(
    bulk.c.start.g = mean((mean.c.pre/100) * soil.start.wt), 
    bulk.c.end.g = mean((mean.c.post/100) * (soil.start.wt - (soil.start.wt*(mean.pct.mass.loss/100)))), 
    frac.mass.start.g = sum(weight.g),
    frac.mass.end.g = sum(weight.g - (weight.g*(pct.mass.loss/100))),
    bulk.pct.mass.loss = mean(mean.pct.mass.loss),
    frac.pct.mass.loss = mean(100 - ((frac.mass.end.g/frac.mass.start.g)*100)),
    frac.c.start.g = sum((pct.c.pre.ashing/100)*weight.g),
    frac.c.end.g = sum((pct.c.post.ashing/100)*(weight.g-(weight.g*(pct.mass.loss/100))))
  )
summary(cmod_dat)

#standardize variables
library(psycho)

z_cmod_dat <- data.frame(
    z_bulk.c.end.g = standardize(cmod_dat$bulk.c.end.g),
    z_bulk.c.start.g = standardize(cmod_dat$bulk.c.start.g),
    z_frac.mass.start.g = standardize(cmod_dat$frac.mass.start.g),
    z_frac.mass.end.g = standardize(cmod_dat$frac.mass.end.g),
    z_frac.pct.mass.loss = standardize(cmod_dat$frac.pct.mass.loss),
    z_frac.c.start.g = standardize(cmod_dat$frac.c.start.g),
    z_frac.c.end.g = standardize(cmod_dat$frac.c.end.g),
    z_bulk.pct.mass.loss = standardize(cmod_dat$bulk.pct.mass.loss),
    z.site = as.factor(cmod_dat$site),
    z.depth = as.factor(cmod_dat$depth) 
)

#check that it worked
summary(z_cmod_dat)
```


Look at the distribution of our output variable (post-ashing sum of fractions C)
```{r }
(hist<-ggplot(z_cmod_dat, aes(x=z_frac.c.end.g)) +
  geom_histogram() +
  theme_classic())
```

##Fit the model (I am using brms here as it is a nice interface between R - which I understand - and Stan, which I do not).

**Set priors**
```{r}
#see what we can set priors for
get_prior(z_frac.c.end.g ~ z_bulk.c.start.g + z_bulk.c.end.g + z_frac.c.start.g + (1|z.depth), data = z_cmod_dat)
#We can set priors for bulk.c.end.g, bulk.c.start.g, frac.c.start.g
#currently there are no priors set for these coefficients
#
prior1 <- c(set_prior("beta(5, 5)", class = "b", coef = "z_bulk.c.start.g"), 
            set_prior("beta(7.5, 5)", class = "b", coef = "z_bulk.c.end.g"), 
            set_prior("beta(7.5, 5)", class = "b", coef = "z_frac.c.start.g"))
```

**Specify model**
```{r}
cmod_brms <- brm(z_frac.c.end.g ~ z_bulk.c.start.g + z_bulk.c.end.g + z_frac.c.start.g + (1|z.depth), data=z_cmod_dat, prior = prior1, iter = 1000, chains = 4, cores = 2)
```

**View output**
```{r }
summary(cmod_brms)
```

**check model fit**
```{r }
model_fit<- z_cmod_dat %>%
   data_grid(z_frac.c.start.g = seq_range(z_frac.c.start.g, n=100), z_bulk.c.end.g = z_bulk.c.end.g, z_bulk.c.start.g=z_bulk.c.start.g, z.depth=z.depth)%>%
   add_predicted_draws(cmod_brms)%>%
   ggplot(aes(x=z_frac.c.start.g, y=z_frac.c.end.g)) +
   stat_lineribbon(aes(y=.prediction), .width=c(.95, .89, .50),
                   alpha=1/2, colour = "black") + 
   geom_point(data=z_cmod_dat, colour = "darkseagreen4", size = 3) +
   scale_fill_brewer(palette="Greys")
model_fit
#check model convergence
plot(cmod_brms, plotfun = "trace")
```

##Posterior Predictive Checks

The compiled rstanarm models include a y_rep variable in the generated quantiles block (posterior distributions). We can use the pp_check function to see how model predictions compare to the raw data.
```{r }
pp_check(cmod_brms, plotfun="stat", stat = "mean")
pp_check(cmod_brms, plotfun = "dens_overlay")
```


##Try on un-standardized data
```{r }
#get priors
get_prior(frac.c.end.g ~ bulk.c.start.g + bulk.c.end.g + frac.c.start.g + (1|depth), data=cmod_dat)

#priors
prior2 <- c(set_prior("beta(0.01, 0.005)", class = "b", coef = "bulk.c.start.g"), 
            set_prior("beta(0.02, 0.015)", class = "b", coef = "bulk.c.end.g"), 
            set_prior("beta(0.02, 0.01)", class = "b", coef = "frac.c.start.g"))

#model
cmod_brms2 <- brm(frac.c.end.g ~ bulk.c.start.g + bulk.c.end.g + frac.c.start.g + (1|depth), data=cmod_dat, family=Gamma(), prior = prior2, iter = 1000, chains = 4, cores = 2)

#model output
summary(cmod_brms2)
```


**check model fit**
```{r }
model_fit2 <- cmod_dat %>%
   data_grid(frac.start = seq_range(frac.c.start.g, n=100))%>%
   add_predicted_draws(cmod_brms2)%>%
   ggplot(aes(x=site, y=frac.c.end.g)) +
   stat_lineribbon(aes(y=.prediction), .width=c(.95, .89, .50),
                   alpha=1/2, colour = "black") + 
   geom_point(data=cmod_dat, colour = "darkseagreen4", size = 3) +
   scale_fill_brewer(palette="Greys")
model_fit2
#check model convergence
plot(cmod_brms2, plotfun = "trace")
```

##Posterior Predictive Checks
```{r }
pp_check(cmod_brms2, plotfun="stat", stat = "mean")
pp_check(cmod_brms2, plotfun = "dens_overlay")
```

Check priors
```{r }
prior_summary(cmod_brms)
```

#Using predicted inorganic C values:

Once we think we have a good constrained prediction for inorganic C values we can 